\newpage
\phantomsection
\label{prf:n-positive}

\begin{remark*}[Return]
\hyperref[cor:n-positive]{Return to Corollary}
\end{remark*}

\begin{theorem*}[Natural Numbers Are Positive]
For every natural number $n$,
\[
n>0.
\]
\end{theorem*}

\begin{proof}
\textbf{Professional Standard Proof.}~\\
Let $F$ be an ordered field.  We prove that every natural number, viewed in
$F$ by the usual embedding, is positive.

By \hyperref[thm:one-positive]{Theorem~\ref*{thm:one-positive}},
\[
1>0.
\]

Now every natural number is obtained by starting from $1$ and repeatedly
adding $1$.  We prove by induction on $n\in\mathbb{N}$ that the
corresponding element of $F$ is positive.

\emph{Base case.}  For $n=1$, the claim is exactly
\[
1>0.
\]

\emph{Inductive step.}  Assume $n>0$ for some natural number $n$.  Since
\[
1>0,
\]
and the positive part of an ordered field is closed under addition, we have
\[
n+1>0.
\]

Therefore, by induction, every natural number $n$ satisfies
\[
n>0.
\]
\end{proof}

\begin{remark*}[Proof structure]
The proof starts from the positivity of the multiplicative identity and then
propagates that positivity through the natural numbers by repeated addition.
The key closure property is that the sum of two positive elements is again
positive.  Since the natural numbers are built by iterating the successor
operation $n\mapsto n+1$, induction transports the positivity of $1$ to all
of $\mathbb{N}$.
\end{remark*}

\begin{remark*}[Dependencies]~\\
\begin{itemize}
  \item \hyperref[thm:one-positive]{Theorem~\ref*{thm:one-positive}}
  \item \hyperref[def:ordered-field]{Definition~\ref*{def:ordered-field}}
        \textnormal{(closure of the positive part under addition)}
  \item Principle of mathematical induction on $\mathbb{N}$
\end{itemize}
\end{remark*}

