\newpage
\phantomsection
\label{prf:ec-zero-numerator}

\begin{remark*}[Return]
\hyperref[thm:ec-zero-numerator]{Return to Theorem}
\end{remark*}

\begin{theorem*}[Zero Class Criterion]
For every $[(a,b)]\in\mathbb{Q}$,
\[
[(a,b)] = [(0,1)]
\quad\Longleftrightarrow\quad
a = 0.
\]
\end{theorem*}

\begin{proof}
\textbf{Professional Standard Proof.}~\\
Let $[(a,b)]\in\mathbb{Q}$. Then $a,b\in\mathbb{Z}$ and $b\neq 0$.

First suppose that
\[
[(a,b)] = [(0,1)].
\]
By equality of equivalence classes,
\[
(a,b)\sim(0,1).
\]
By the definition of the fraction-pair relation,
\[
a\cdot 1=b\cdot 0.
\]
Since
\[
a\cdot 1=a
\qquad\text{and}\qquad
b\cdot 0=0,
\]
it follows that
\[
a=0.
\]

Conversely, suppose that
\[
a=0.
\]
Then
\[
a\cdot 1=0
\qquad\text{and}\qquad
b\cdot 0=0.
\]
Hence
\[
a\cdot 1=b\cdot 0.
\]
By the definition of the fraction-pair relation,
\[
(a,b)\sim(0,1).
\]
Therefore
\[
[(a,b)] = [(0,1)].
\]

Thus
\[
[(a,b)] = [(0,1)]
\quad\Longleftrightarrow\quad
a=0.
\]
\end{proof}

\begin{proof}
\textbf{Detailed Learning Proof.}~\\

\textbf{Step 1.} Translate the class equality into the representative relation.

Let $[(a,b)]\in\mathbb{Q}$. Then $(a,b)$ is an admissible fraction-pair,
so $a,b\in\mathbb{Z}$ and $b\neq 0$.

The equation
\[
[(a,b)] = [(0,1)]
\]
means exactly that the representatives $(a,b)$ and $(0,1)$ are equivalent:
\[
(a,b)\sim(0,1).
\]

\textbf{Step 2.} Prove the forward implication.

Assume
\[
[(a,b)] = [(0,1)].
\]
Then
\[
(a,b)\sim(0,1).
\]
By the definition of the fraction-pair relation,
\[
a\cdot 1=b\cdot 0.
\]
Using the integer identity law and zero law,
\[
a\cdot 1=a
\qquad\text{and}\qquad
b\cdot 0=0.
\]
Therefore
\[
a=0.
\]

\textbf{Step 3.} Prove the reverse implication.

Assume
\[
a=0.
\]
Then
\[
a\cdot 1=0
\]
and
\[
b\cdot 0=0.
\]
Hence
\[
a\cdot 1=b\cdot 0.
\]
By the definition of the fraction-pair relation,
\[
(a,b)\sim(0,1).
\]
Therefore the corresponding equivalence classes are equal:
\[
[(a,b)] = [(0,1)].
\]

\textbf{Step 4.} Conclude the biconditional.

The forward implication gives
\[
[(a,b)] = [(0,1)] \;\Longrightarrow\; a=0.
\]
The reverse implication gives
\[
a=0 \;\Longrightarrow\; [(a,b)] = [(0,1)].
\]
Therefore
\[
[(a,b)] = [(0,1)]
\quad\Longleftrightarrow\quad
a=0.
\]
\end{proof}

\begin{remark*}[Proof structure]
The proof is a direct proof of both implications. The class equality is
first translated into the defining equivalence relation on fraction-pairs.
The resulting cross-product equality is
\[
a\cdot 1=b\cdot 0,
\]
which reduces to $a=0$ by elementary integer arithmetic.
\end{remark*}

\begin{remark*}[Dependencies]~\\
\begin{itemize}
  \item \hyperref[def:fraction-pair-equality]{Definition~\ref*{def:fraction-pair-equality}}
  \item \hyperref[def:rational-distinguished]{Definition~\ref*{def:rational-distinguished}}
        \textnormal{($0 = [(0,1)]$)}
  \item Basic integer arithmetic (cancellation, zero product)
\end{itemize}
\end{remark*}