\newpage
\phantomsection
\label{prf:product-of-negatives}

\begin{remark*}[Return]
\hyperref[cor:product-of-negatives]{Return to Corollary}
\end{remark*}

\begin{corollary*}[Product of Negatives]
Let $F$ be a field and let $a,b\in F$.  Then
\[
(-a)\cdot(-b) = a\cdot b.
\]
\end{corollary*}

\begin{proof}
\textbf{Professional Standard Proof.}~\\
By the negation of a product theorem applied twice:
\[
(-a)\cdot(-b) = -(a\cdot(-b)) = -(-(a\cdot b)) = a\cdot b.
\]
\end{proof}

\begin{proof}
\textbf{Detailed Learning Proof.}~\\

\textbf{Step 1.} Apply the negation of a product theorem to the outer negation.

By \hyperref[thm:negation-of-product]{Theorem~\ref*{thm:negation-of-product}}
with $a$ replaced by $a$ and $b$ replaced by $-b$:
\[
(-a)\cdot(-b) = -(a\cdot(-b)).
\]

\textbf{Step 2.} Apply the negation of a product theorem to the inner term.

By \hyperref[thm:negation-of-product]{Theorem~\ref*{thm:negation-of-product}},
\[
a\cdot(-b) = -(a\cdot b).
\]
Substituting into Step 1:
\[
(-a)\cdot(-b) = -(-(a\cdot b)).
\]

\textbf{Step 3.} Apply double negation.

By \hyperref[thm:double-negation]{Theorem~\ref*{thm:double-negation}},
\[
-(-(a\cdot b)) = a\cdot b.
\]

\textbf{Step 4.} Conclude.

Chaining Steps 1--3:
\[
(-a)\cdot(-b) = -(-(a\cdot b)) = a\cdot b. \qedhere
\]
\end{proof}

\begin{remark*}[Proof structure]
This corollary is a two-line chain from the negation of a product theorem
and double negation.  The negation of a product theorem is applied twice in
sequence: first to pull $(-a)$ out of the product, producing $-(a\cdot(-b))$,
then to pull $(-b)$ out, producing $-(-(a\cdot b))$.  Double negation then
collapses the two negation signs to nothing.  No field axioms are cited
directly; all work was done in the two preceding theorems.
\end{remark*}

\begin{remark*}[Dependencies]~\\
\begin{itemize}
  \item \hyperref[thm:negation-of-product]{Theorem~\ref*{thm:negation-of-product}}
        \textnormal{($(-a)\cdot b = -(a\cdot b)$ and $a\cdot(-b) = -(a\cdot b)$)}
  \item \hyperref[thm:double-negation]{Theorem~\ref*{thm:double-negation}}
        \textnormal{($-(-(a\cdot b)) = a\cdot b$)}
\end{itemize}
\end{remark*}
