\newpage
\phantomsection
\label{prf:reciprocal-order-for-positive-elements}

\begin{remark*}[Return]
\hyperref[thm:reciprocal-order-for-positive-elements]{Return to Theorem}
\end{remark*}

\begin{theorem*}[Reciprocal Order for Positive Elements]
If $0<d<b$, then
\[
0<\frac{1}{b}<\frac{1}{d}.
\]
Conversely, if
\[
0<\frac{1}{b}<\frac{1}{d},
\]
then $0<d<b$.
\end{theorem*}

\begin{proof}
\textbf{Professional Standard Proof.}~\\
\emph{Forward direction.} Assume
\[
0<d<b.
\]
Then in particular
\[
d>0
\qquad\text{and}\qquad
b>0.
\]
By \hyperref[thm:inverse-positive]{Theorem~\ref*{thm:inverse-positive}},
the inverses of positive elements are positive.  Hence
\[
0<\frac{1}{d}
\qquad\text{and}\qquad
0<\frac{1}{b}.
\]

It remains to compare the two reciprocals.  Since $b>0$ and $d>0$, we have
\[
0<b\cdot d.
\]
Now apply
\hyperref[thm:fraction-order-comparison-positive-denominator-product]{Theorem~\ref*{thm:fraction-order-comparison-positive-denominator-product}}
with numerators $1$ and $1$:
\[
\frac{1}{b}<\frac{1}{d}
\quad\Longleftrightarrow\quad
1\cdot d<b\cdot 1.
\]
But
\[
1\cdot d=d
\qquad\text{and}\qquad
b\cdot 1=b,
\]
so the right-hand inequality is exactly
\[
d<b,
\]
which holds by hypothesis.  Therefore
\[
\frac{1}{b}<\frac{1}{d}.
\]
Combining this with $0<1/b$, we obtain
\[
0<\frac{1}{b}<\frac{1}{d}.
\]

\emph{Converse direction.} Assume
\[
0<\frac{1}{b}<\frac{1}{d}.
\]
From
\[
0<\frac{1}{b},
\]
\hyperref[thm:positive-fraction-iff-positive-product]{Theorem~\ref*{thm:positive-fraction-iff-positive-product}}
with numerator $1$ gives
\[
0<1\cdot b=b.
\]
So $b>0$.  Similarly, from
\[
0<\frac{1}{d},
\]
the same theorem gives
\[
0<1\cdot d=d,
\]
so $d>0$.

Thus
\[
0<b\cdot d.
\]
Applying
\hyperref[thm:fraction-order-comparison-positive-denominator-product]{Theorem~\ref*{thm:fraction-order-comparison-positive-denominator-product}}
again with numerators $1$ and $1$, we obtain
\[
\frac{1}{b}<\frac{1}{d}
\quad\Longleftrightarrow\quad
1\cdot d<b\cdot 1.
\]
Hence
\[
d<b.
\]
Together with $d>0$, this yields
\[
0<d<b.
\]
\end{proof}

\begin{remark*}[Proof structure]
The proof isolates two separate issues: positivity of the reciprocals and
their relative order.  Positivity comes from the inverse-positive theorem in
the forward direction, and from the quotient-positivity criterion in the
reverse direction.  Once positivity of the denominators is known, the actual
comparison is just the positive-denominator fraction comparison theorem
applied to the special fractions $1/b$ and $1/d$.  In that specialization,
the cross-multiplied inequality collapses exactly to $d<b$.
\end{remark*}

\begin{remark*}[Dependencies]~\\
\begin{itemize}
  \item \hyperref[thm:fraction-order-comparison-positive-denominator-product]{Theorem~\ref*{thm:fraction-order-comparison-positive-denominator-product}}
  \item \hyperref[thm:inverse-positive]{Theorem~\ref*{thm:inverse-positive}}
  \item \hyperref[thm:positive-fraction-iff-positive-product]{Theorem~\ref*{thm:positive-fraction-iff-positive-product}}
\end{itemize}
\end{remark*}
