% =========================================================
% Proof: Mediant Gap Lemma
% =========================================================

\newpage
\phantomsection
\label{prf:mediant-gap}

\begin{remark*}[Return]
\hyperref[lem:mediant-gap]{Return to Lemma~\ref*{lem:mediant-gap}.}
\end{remark*}

\begin{theorem*}[Mediant Gap Lemma]
If $a/b < c/d$ with $b, d > 0$, then
\[
\frac{a+c}{b+d} - \frac{a}{b} = \frac{1}{b(b+d)},
\qquad
\frac{c}{d} - \frac{a+c}{b+d} = \frac{1}{d(b+d)}.
\]
\end{theorem*}

\begin{proof}[Proof (Professional Standard)]
\textbf{Left gap.}
\[
\frac{a+c}{b+d} - \frac{a}{b}
= \frac{b(a+c) - a(b+d)}{b(b+d)}
= \frac{ba + bc - ab - ad}{b(b+d)}
= \frac{bc - ad}{b(b+d)}.
\]
Since $a/b < c/d$ with $b,d > 0$, cross-multiplication gives $ad < bc$,
so $bc - ad > 0$.  For the Farey setting where $a/b$ and $c/d$ are
neighbors, $bc - ad = 1$ by the Farey Neighbor Theorem; in the general
case (any $a/b < c/d$ with positive denominators) one has $bc - ad \geq 1$.
The formula gives the exact value in terms of $bc - ad$.

When $bc - ad = 1$ (the Farey case):
\[
\frac{a+c}{b+d} - \frac{a}{b} = \frac{1}{b(b+d)}.
\]

\textbf{Right gap.}
\[
\frac{c}{d} - \frac{a+c}{b+d}
= \frac{c(b+d) - d(a+c)}{d(b+d)}
= \frac{cb + cd - da - dc}{d(b+d)}
= \frac{bc - ad}{d(b+d)}.
\]
When $bc - ad = 1$:
\[
\frac{c}{d} - \frac{a+c}{b+d} = \frac{1}{d(b+d)}.
\qedhere
\]
\end{proof}

\begin{proof}[Proof (Detailed Learning)]
\textbf{Step 1: Compute the left gap by combining over a common denominator.}
\[
\frac{a+c}{b+d} - \frac{a}{b}
= \frac{b(a+c)}{b(b+d)} - \frac{a(b+d)}{b(b+d)}
= \frac{b(a+c) - a(b+d)}{b(b+d)}.
\]

\textbf{Step 2: Expand and simplify the numerator.}
\[
b(a+c) - a(b+d)
= ba + bc - ab - ad
= bc - ad.
\]
So
\[
\frac{a+c}{b+d} - \frac{a}{b} = \frac{bc - ad}{b(b+d)}.
\]

\textbf{Step 3: Apply the Farey Neighbor Theorem.}
Since $a/b$ and $c/d$ are Farey neighbors with $\gcd(a,b) = \gcd(c,d) = 1$,
the \hyperref[thm:farey-neighbor-theorem]{Farey Neighbor Theorem} gives
$bc - ad = 1$.  Therefore:
\[
\frac{a+c}{b+d} - \frac{a}{b} = \frac{1}{b(b+d)}.
\]

\textbf{Step 4: Compute the right gap symmetrically.}
\[
\frac{c}{d} - \frac{a+c}{b+d}
= \frac{c(b+d)}{d(b+d)} - \frac{d(a+c)}{d(b+d)}
= \frac{c(b+d) - d(a+c)}{d(b+d)}.
\]
Expand:
\[
c(b+d) - d(a+c)
= cb + cd - da - dc
= bc - ad = 1.
\]
Therefore:
\[
\frac{c}{d} - \frac{a+c}{b+d} = \frac{1}{d(b+d)}.
\]

\textbf{Step 5: Verify the subgaps are smaller than the original gap.}
The original gap is $c/d - a/b = 1/(bd)$ (by the
\hyperref[thm:farey-gap]{gap formula}).
Since $b + d > b$ and $b + d > d$:
\[
\frac{1}{b(b+d)} < \frac{1}{b \cdot b} = \frac{1}{b^2}
\quad\text{and}\quad
\frac{1}{d(b+d)} < \frac{1}{d^2}.
\]
Both subgaps are strictly smaller than the original gap $1/(bd)$, confirming
that mediant insertion always refines the interval.
\end{proof}

\begin{remark*}[Proof structure]
Two symmetric computations: subtract over a common denominator, expand and
cancel, substitute $bc - ad = 1$ from the Farey Neighbor Theorem.  The
computation is elementary; the significance is in what the formulas imply
for the shrinkage rate of Stern--Brocot intervals.
\end{remark*}

\begin{remark*}[Dependencies]
\hyperref[thm:farey-neighbor-theorem]{Farey Neighbor Theorem},
\hyperref[thm:mediant-inequality]{Mediant inequality},
fraction arithmetic in $\mathbb{Q}$.
\end{remark*}
