\newpage
\phantomsection
\label{prf:ec-subtraction}

\begin{remark*}[Return]
\hyperref[thm:ec-subtraction]{Return to Theorem}
\end{remark*}

\begin{theorem*}[Subtraction of Equivalence Classes]
For every $[(a,b)],[(c,d)]\in\mathbb{Q}$,
\[
[(a,b)] - [(c,d)] = [(ad-bc,\,bd)].
\]
\end{theorem*}

\begin{proof}
\textbf{Professional Standard Proof.}~\\
By definition, $[(a,b)] - [(c,d)] := [(a,b)] + (-[(c,d)])$.

By the negation theorem, $-[(c,d)] = [(-c,d)]$.

By the addition formula,
\[
[(a,b)] + [(-c,d)]
= [(a\cdot d + b\cdot(-c),\; b\cdot d)]
= [(ad - bc,\; bd)].
\]
Therefore $[(a,b)] - [(c,d)] = [(ad-bc,bd)]$.
\end{proof}

\begin{proof}
\textbf{Detailed Learning Proof.}~\\

\textbf{Step 1.} Recall the definition of subtraction.

Subtraction in any group is defined as addition of the additive inverse:
\[
[(a,b)] - [(c,d)] := [(a,b)] + \bigl(-[(c,d)]\bigr).
\]

\textbf{Step 2.} Apply the negation formula.

By the negation theorem,
\[
-[(c,d)] = [(-c,d)].
\]
This is valid since $(c,d)\in W$ implies $d\neq 0$, and $-c\in\mathbb{Z}$,
so $(-c,d)\in W$.

\textbf{Step 3.} Apply the addition formula.

By the addition theorem,
\[
[(a,b)] + [(-c,d)]
= [(a\cdot d + b\cdot(-c),\; b\cdot d)].
\]

\textbf{Step 4.} Simplify the numerator.

The numerator is
\[
a\cdot d + b\cdot(-c) = ad - bc.
\]

\textbf{Step 5.} Verify the denominator is nonzero.

Since $b\neq 0$ and $d\neq 0$, we have $bd\neq 0$, so $(ad-bc,bd)\in W$
and $[(ad-bc,bd)]\in\mathbb{Q}$.

\textbf{Step 6.} Assemble the conclusion.

Combining Steps 1--5:
\[
[(a,b)] - [(c,d)]
= [(a,b)] + [(-c,d)]
= [(ad-bc,\; bd)]. \qedhere
\]
\end{proof}

\begin{remark*}[Proof structure]
The proof is a chain of definitions. Subtraction is addition of the
negation; the negation formula swaps the sign of the numerator; the
addition formula then produces the stated result. The only integer
arithmetic is $b(-c) = -bc$. This theorem depends on both the negation
and addition theorems, which is why it appears after them.
\end{remark*}

\begin{remark*}[Dependencies]~\\
\begin{itemize}
  \item Definition of subtraction as $x - y := x + (-y)$
  \item \hyperref[thm:ec-negation]{Theorem~\ref*{thm:ec-negation}}
        \textnormal{($-[(c,d)] = [(-c,d)]$)}
  \item \hyperref[thm:ec-addition]{Theorem~\ref*{thm:ec-addition}}
        \textnormal{(addition formula for equivalence classes)}
  \item Basic integer arithmetic ($b(-c) = -bc$)
\end{itemize}
\end{remark*}
