% =========================================================
% Proof: Convergent Sequences Are Bounded
% =========================================================

\newpage
\phantomsection
\label{prf:convergent-bounded-q}

\begin{remark*}[Return]
\hyperref[thm:convergent-bounded-q]{Return to Theorem~\ref*{thm:convergent-bounded-q}.}
\end{remark*}

\begin{theorem*}[Convergent sequences are bounded]
Every convergent sequence of rational numbers is bounded.
\end{theorem*}

\begin{proof}[Proof (Professional Standard)]
Let $(x_n)$ be a sequence of rational numbers with $x_n \to a \in \mathbb{Q}$.
Apply the definition of convergence with $\varepsilon = 1$: there exists
$N \in \mathbb{N}$ such that $|x_n - a| < 1$ for all $n \geq N$.
By the triangle inequality, $|x_n| \leq |x_n - a| + |a| < 1 + |a|$
for all $n \geq N$.
Define
\[
\Gamma = \max\{|x_1|, |x_2|, \ldots, |x_N|, 1 + |a|\}.
\]
Then $|x_n| \leq \Gamma$ for all $n \in \mathbb{N}$, so $(x_n)$ is bounded.
\end{proof}

\begin{proof}[Proof (Detailed Learning)]
\textbf{Step 1: Control the tail using convergence.}
Since $x_n \to a$, choose $\varepsilon = 1$.  There exists $N \in \mathbb{N}$
such that $|x_n - a| < 1$ for all $n \geq N$.

\textbf{Step 2: Bound each tail term.}
For any $n \geq N$, apply the triangle inequality:
\[
|x_n| = |x_n - a + a| \leq |x_n - a| + |a| < 1 + |a|.
\]

\textbf{Step 3: Handle the initial segment.}
The terms $x_1, x_2, \ldots, x_N$ are finitely many rational numbers,
each with a finite absolute value.  Their maximum $\max\{|x_1|,\ldots,|x_N|\}$
is a well-defined rational number.

\textbf{Step 4: Combine into a single bound.}
Set
\[
\Gamma = \max\bigl\{|x_1|, |x_2|, \ldots, |x_N|,\; 1 + |a|\bigr\}.
\]
For $n \leq N$: $|x_n| \leq \max\{|x_1|,\ldots,|x_N|\} \leq \Gamma$.
For $n > N$: $|x_n| < 1 + |a| \leq \Gamma$.
Therefore $|x_n| \leq \Gamma$ for all $n \in \mathbb{N}$.

\textbf{Step 5: Conclude.}
By \hyperref[def:bounded-sequence-q]{Definition~\ref*{def:bounded-sequence-q}},
$(x_n)$ is bounded.
\end{proof}

\begin{remark*}[Proof structure]
The proof splits the sequence into two parts: a finite initial segment
(handled by taking a maximum of finitely many values) and a tail
(controlled by convergence with $\varepsilon = 1$).  The choice
$\varepsilon = 1$ is not special — any positive $\varepsilon$ works — but
$1$ keeps the arithmetic clean.
\end{remark*}

\begin{remark*}[Dependencies]
\hyperref[def:convergent-rational-sequence]{Convergence of a rational sequence},
\hyperref[def:bounded-sequence-q]{Bounded sequence},
the triangle inequality in $\mathbb{Q}$.
\end{remark*}
