\newpage
\phantomsection
\label{prf:ec-integer-embedding-mul}

\begin{remark*}[Return]
\hyperref[thm:ec-integer-embedding-mul]{Return to Theorem}
\end{remark*}

\begin{theorem*}[Integer Embedding Preserves Multiplication]
For all $m,n\in\mathbb{Z}$,
\[
[(m,1)] \cdot [(n,1)] = [(mn,\,1)].
\]
\end{theorem*}

\begin{proof}
\textbf{Professional Standard Proof.}~\\
Let $m,n\in\mathbb{Z}$.  By the definition of multiplication on rational
equivalence classes,
\[
[(m,1)] \cdot [(n,1)]
= [(m\cdot n,\; 1\cdot 1)]
= [(mn,\; 1)].
\]
\end{proof}

\begin{proof}
\textbf{Detailed Learning Proof.}~\\

\textbf{Step 1.} Identify the embeddings.

Let $m,n\in\mathbb{Z}$.  Under the integer embedding, $m$ is identified
with $[(m,1)]\in\mathbb{Q}$ and $n$ is identified with $[(n,1)]\in\mathbb{Q}$.
Both pairs $(m,1)$ and $(n,1)$ are admissible fraction-pairs since $1\neq 0$.

\textbf{Step 2.} Apply the multiplication formula.

By the definition of multiplication on rational equivalence classes,
\[
[(m,1)] \cdot [(n,1)]
:= [(m\cdot n,\; 1\cdot 1)].
\]

\textbf{Step 3.} Simplify using integer arithmetic.

In $\mathbb{Z}$:
\[
m\cdot n = mn
\qquad\text{and}\qquad
1\cdot 1 = 1.
\]
Therefore
\[
[(m\cdot n,\; 1\cdot 1)] = [(mn,\; 1)].
\]

\textbf{Step 4.} Conclude.

Combining Steps 2 and 3:
\[
[(m,1)] \cdot [(n,1)] = [(mn,\,1)].
\]
This is the embedding of the integer $mn$, so the embedding preserves
multiplication.
\end{proof}

\begin{remark*}[Proof structure]
The proof is structurally identical to the addition embedding proof: apply
the operation formula at the specific input $(m,1)$, $(n,1)$ and simplify
the resulting integer expressions.  For multiplication the formula gives
numerator $m\cdot n$ and denominator $1\cdot 1$, both of which collapse
immediately.  No equivalence relation work is needed; well-definedness was
discharged by the well-definedness lemma at the point of definition.
\end{remark*}

\begin{remark*}[Dependencies]~\\
\begin{itemize}
  \item \hyperref[def:rational-operations]{Definition~\ref*{def:rational-operations}}
        \textnormal{(multiplication formula: $[(a,b)]\cdot[(c,d)]:=[(ac,bd)]$)}
  \item \hyperref[lem:rational-multiplication-well-defined]{Lemma~\ref*{lem:rational-multiplication-well-defined}}
        \textnormal{(well-definedness of multiplication, discharges representative independence)}
  \item Basic integer arithmetic ($m\cdot n = mn$, $1\cdot 1 = 1$)
\end{itemize}
\end{remark*}
