\newpage
\phantomsection
\label{prf:negative-exponents-agree-with-inverse-power}

\begin{remark*}[Return]
\hyperref[thm:negative-exponents-agree-with-inverse-power]{Return to Theorem}
\end{remark*}

\begin{theorem*}[Negative Exponents Agree with Inverse Powers]
Let $F$ be a field, let $x\in F$ with $x\neq 0$, and let $m\in\mathbb{N}$.
Then
\[
x^{-m} = (x^{-1})^m = (x^m)^{-1}.
\]
\end{theorem*}

\begin{proof}
\textbf{Professional Standard Proof.}~\\
The first equality $x^{-m} = (x^{-1})^m$ is the definition of negative
integer exponents.

For $x^{-m} = (x^m)^{-1}$, we proceed by induction on $m$.

\emph{Base case} $m=0$: $x^0 = 1$ and $(x^0)^{-1} = 1^{-1} = 1$, so
$x^{-0} = 1 = (x^0)^{-1}$.

\emph{Inductive step}: Assume $(x^k)^{-1} = x^{-k}$ for some $k\in\mathbb{N}$.
Then
\[
(x^{k+1})^{-1}
= (x^k\cdot x)^{-1}
= x^{-1}\cdot (x^k)^{-1}
= x^{-1}\cdot x^{-k}
= (x^{-1})^1\cdot(x^{-1})^k
= (x^{-1})^{k+1}
= x^{-(k+1)}.
\]
By induction, $(x^m)^{-1} = x^{-m}$ for all $m\in\mathbb{N}$.
\end{proof}

\begin{proof}
\textbf{Detailed Learning Proof.}~\\

We prove the two equalities separately.

\textbf{Part A: $x^{-m} = (x^{-1})^m$.}

This is immediate from the definition of negative integer exponents
(\hyperref[def:negative-integer-exponents]{Definition~\ref*{def:negative-integer-exponents}}),
which states:
\[
x^{-m} := (x^{-1})^m
\]
for $x\neq 0$ and $m\in\mathbb{N}$.  No computation is required; this equality
holds by definition.

\textbf{Part B: $x^{-m} = (x^m)^{-1}$.}

We prove $(x^m)^{-1} = (x^{-1})^m$ by induction on $m\in\mathbb{N}$, which
combined with Part A gives $x^{-m} = (x^m)^{-1}$.

\textbf{Base case} $m = 0$:

By the nonnegative exponent definition, $x^0 = 1$.  Therefore
$(x^0)^{-1} = 1^{-1} = 1$ by \hyperref[thm:one-inverse-is-one]{Theorem~\ref*{thm:one-inverse-is-one}}.
Also $(x^{-1})^0 = 1$ by the nonnegative exponent definition applied to $x^{-1}$.
So $(x^0)^{-1} = 1 = (x^{-1})^0$, establishing the base case.

\textbf{Inductive step}:

Assume the statement holds for some $k\in\mathbb{N}$, i.e.\ $(x^k)^{-1} = (x^{-1})^k$.
We must show $(x^{k+1})^{-1} = (x^{-1})^{k+1}$.

\textbf{Step B1.} Expand $x^{k+1}$ using the recursive definition.

By \hyperref[def:nonneg-integer-exponents-field]{Definition~\ref*{def:nonneg-integer-exponents-field}},
\[
x^{k+1} = x^k \cdot x.
\]

\textbf{Step B2.} Invert the product.

By the field theorem on the inverse of a product (the inverse reverses order:
$(a\cdot b)^{-1} = b^{-1}\cdot a^{-1}$, which equals $a^{-1}\cdot b^{-1}$
by commutativity of multiplication),
\[
(x^{k+1})^{-1} = (x^k\cdot x)^{-1} = x^{-1}\cdot (x^k)^{-1}.
\]

\textbf{Step B3.} Apply the inductive hypothesis.

By the inductive hypothesis, $(x^k)^{-1} = (x^{-1})^k$.  Substituting:
\[
x^{-1}\cdot(x^k)^{-1} = x^{-1}\cdot(x^{-1})^k.
\]

\textbf{Step B4.} Collapse using the recursive definition applied to $x^{-1}$.

By \hyperref[def:nonneg-integer-exponents-field]{Definition~\ref*{def:nonneg-integer-exponents-field}}
applied to the base element $x^{-1}$:
\[
(x^{-1})^{k+1} = (x^{-1})^k \cdot x^{-1} = x^{-1}\cdot(x^{-1})^k.
\]
(using commutativity of multiplication for the last step).  Therefore:
\[
(x^{k+1})^{-1} = (x^{-1})^{k+1}.
\]

\textbf{Step B5.} Close the induction.

By the principle of mathematical induction, $(x^m)^{-1} = (x^{-1})^m$ for
all $m\in\mathbb{N}$.  Combined with Part A:
\[
x^{-m} = (x^{-1})^m = (x^m)^{-1}. \qedhere
\]
\end{proof}

\begin{remark*}[Proof structure]
The first equality is definitional.  The second requires induction because
it asks about the relationship between two independently defined operations:
natural-number exponentiation $x^m$ (built by repeated multiplication) and
the field inverse $(-)^{-1}$ (an axiomatically given operation).  The inductive
step turns on the product-inversion identity $(ab)^{-1} = b^{-1}a^{-1}$, which
lets the inversion distribute through one factor at a time, consuming $x^k$
via the inductive hypothesis.  The base case $m=0$ requires
\hyperref[thm:one-inverse-is-one]{Theorem~\ref*{thm:one-inverse-is-one}} to
identify $1^{-1} = 1$.
\end{remark*}

\begin{remark*}[Dependencies]~\\
\begin{itemize}
  \item \hyperref[def:negative-integer-exponents]{Definition~\ref*{def:negative-integer-exponents}}
        \textnormal{($x^{-m} := (x^{-1})^m$; this is the first equality)}
  \item \hyperref[def:nonneg-integer-exponents-field]{Definition~\ref*{def:nonneg-integer-exponents-field}}
        \textnormal{(recursive definition $x^{k+1} = x^k\cdot x$, used in inductive step)}
  \item \hyperref[def:field]{Definition~\ref*{def:field}}
        \textnormal{(multiplicative inverse; commutativity of multiplication)}
  \item \hyperref[thm:one-inverse-is-one]{Theorem~\ref*{thm:one-inverse-is-one}}
        \textnormal{($1^{-1} = 1$; used in base case)}
  \item \hyperref[thm:uniqueness-of-multiplicative-inverse]{Theorem~\ref*{thm:uniqueness-of-multiplicative-inverse}}
        \textnormal{(inverse of a product: $(ab)^{-1} = b^{-1}a^{-1}$)}
  \item Principle of mathematical induction on $\mathbb{N}$
\end{itemize}
\end{remark*}
