\newpage
\phantomsection
\label{prf:rational-cross-multiplication}

\begin{remark*}[Return]
\hyperref[thm:rational-cross-multiplication]{Return to Theorem}
\end{remark*}

\begin{theorem*}[Cross-Multiplication Criterion for Rational Equality]
For every $a,b,c,d\in\mathbb{Q}$ with $b\neq 0$ and $d\neq 0$,
\[
\frac{a}{b}=\frac{c}{d}
\quad\Longleftrightarrow\quad
a\cdot d=b\cdot c.
\]
\end{theorem*}

\begin{proof}
\textbf{Professional Standard Proof.}~\\
Let $a,b,c,d\in\mathbb{Q}$ with $b\neq 0$ and $d\neq 0$.

First suppose that
\[
\frac{a}{b}=\frac{c}{d}.
\]
By the definition of rational division,
\[
a\cdot b^{-1}=c\cdot d^{-1}.
\]
Multiplying both sides by $bd$ gives
\[
(a\cdot b^{-1})\cdot bd=(c\cdot d^{-1})\cdot bd.
\]
Using associativity, commutativity, and the inverse laws in $\mathbb{Q}$,
this simplifies to
\[
a\cdot d=b\cdot c.
\]

Conversely, suppose that
\[
a\cdot d=b\cdot c.
\]
Since $b\neq 0$ and $d\neq 0$, the product $bd$ is nonzero. Multiplying
both sides by $(bd)^{-1}$ gives
\[
(a\cdot d)(bd)^{-1}=(b\cdot c)(bd)^{-1}.
\]
Using associativity, commutativity, and the inverse laws in $\mathbb{Q}$,
this simplifies to
\[
a\cdot b^{-1}=c\cdot d^{-1}.
\]
By the definition of rational division,
\[
\frac{a}{b}=\frac{c}{d}.
\]
Therefore
\[
\frac{a}{b}=\frac{c}{d}
\quad\Longleftrightarrow\quad
a\cdot d=b\cdot c.
\]
\end{proof}

\begin{proof}
\textbf{Detailed Learning Proof.}~\\

\textbf{Step 1.} Rewrite the quotients using multiplicative inverses.

Let $a,b,c,d\in\mathbb{Q}$ with
\[
b\neq 0
\qquad\text{and}\qquad
d\neq 0.
\]
By the definition of rational division,
\[
\frac{a}{b}=a\cdot b^{-1}
\qquad\text{and}\qquad
\frac{c}{d}=c\cdot d^{-1}.
\]

\textbf{Step 2.} Prove the forward implication.

Assume
\[
\frac{a}{b}=\frac{c}{d}.
\]
Then
\[
a\cdot b^{-1}=c\cdot d^{-1}.
\]
Multiplying both sides by $bd$ gives
\[
(a\cdot b^{-1})\cdot bd=(c\cdot d^{-1})\cdot bd.
\]
The left-hand side simplifies as follows:
\[
(a\cdot b^{-1})\cdot bd
=
a\cdot (b^{-1}\cdot b)\cdot d
=
a\cdot 1\cdot d
=
a\cdot d.
\]
The right-hand side simplifies as follows:
\[
(c\cdot d^{-1})\cdot bd
=
c\cdot b\cdot (d^{-1}\cdot d)
=
c\cdot b\cdot 1
=
b\cdot c.
\]
Therefore
\[
a\cdot d=b\cdot c.
\]

\textbf{Step 3.} Prove the reverse implication.

Assume
\[
a\cdot d=b\cdot c.
\]
Since $b\neq 0$ and $d\neq 0$, the product $bd$ is nonzero. Therefore
$(bd)^{-1}$ exists in $\mathbb{Q}$. Multiplying both sides by $(bd)^{-1}$
gives
\[
(a\cdot d)(bd)^{-1}=(b\cdot c)(bd)^{-1}.
\]
Since $\mathbb{Q}$ is a field,
\[
(bd)^{-1}=b^{-1}d^{-1}.
\]
Hence the left-hand side becomes
\[
(a\cdot d)(bd)^{-1}
=
a\cdot d\cdot b^{-1}\cdot d^{-1}
=
a\cdot b^{-1}\cdot (d\cdot d^{-1})
=
a\cdot b^{-1}.
\]
Similarly, the right-hand side becomes
\[
(b\cdot c)(bd)^{-1}
=
b\cdot c\cdot b^{-1}\cdot d^{-1}
=
c\cdot (b\cdot b^{-1})\cdot d^{-1}
=
c\cdot d^{-1}.
\]
Therefore
\[
a\cdot b^{-1}=c\cdot d^{-1}.
\]
By the definition of rational division,
\[
\frac{a}{b}=\frac{c}{d}.
\]

\textbf{Step 4.} Conclude the biconditional.

The forward implication proves
\[
\frac{a}{b}=\frac{c}{d}
\;\Longrightarrow\;
a\cdot d=b\cdot c.
\]
The reverse implication proves
\[
a\cdot d=b\cdot c
\;\Longrightarrow\;
\frac{a}{b}=\frac{c}{d}.
\]
Therefore
\[
\frac{a}{b}=\frac{c}{d}
\quad\Longleftrightarrow\quad
a\cdot d=b\cdot c.
\]
\end{proof}

\begin{remark*}[Proof structure]
The proof is a direct proof of both implications. The forward direction
rewrites rational division as multiplication by inverses and then clears
denominators by multiplying by $bd$. The reverse direction starts from
the cross-product equality and multiplies by the inverse of the nonzero
product $bd$ to recover equality of the two quotients.
\end{remark*}

\begin{remark*}[Dependencies]~\\
\begin{itemize}
  \item \hyperref[def:rational-division]{Definition~\ref*{def:rational-division}}
  \item \hyperref[thm:q-is-integral-domain]{Theorem~\ref*{thm:q-is-integral-domain}}
  \item \hyperref[thm:q-is-field]{Theorem~\ref*{thm:q-is-field}}
  \item Associativity and commutativity of multiplication in $\mathbb{Q}$
  \item Multiplicative inverse law in $\mathbb{Q}$
  \item Nonzero product property in $\mathbb{Q}$
\end{itemize}
\end{remark*}