% =========================================================
% Proof: The Cauchy Equivalence Relation on Q^c
% =========================================================

\newpage
\phantomsection
\label{prf:cauchy-equivalence-is-equivalence}

\begin{remark*}[Return]
\hyperref[thm:cauchy-equivalence-is-equivalence]{Return to Theorem~\ref*{thm:cauchy-equivalence-is-equivalence}.}
\end{remark*}

\begin{theorem*}[The relation $\sim$ is an equivalence relation on $\mathbb{Q}^c$]
Define $(x_n) \sim (y_n)$ if and only if $\lim_{n\to\infty}|x_n - y_n| = 0$.
Then $\sim$ is an equivalence relation on the set $\mathbb{Q}^c$ of all
Cauchy sequences in $\mathbb{Q}$.
\end{theorem*}

\begin{proof}[Proof (Professional Standard)]
\textbf{Reflexivity.}
For any $(x_n) \in \mathbb{Q}^c$: $|x_n - x_n| = 0 < \varepsilon$ for all
$\varepsilon > 0$ and all $n$, so $(x_n) \sim (x_n)$.

\textbf{Symmetry.}
If $(x_n) \sim (y_n)$ then $|x_n - y_n| \to 0$.
Since $|y_n - x_n| = |x_n - y_n|$, we have $|y_n - x_n| \to 0$, so
$(y_n) \sim (x_n)$.

\textbf{Transitivity.}
Suppose $(x_n) \sim (y_n)$ and $(y_n) \sim (z_n)$.
Let $\varepsilon > 0$.  There exist $N_1, N_2 \in \mathbb{N}$ such that
$|x_n - y_n| < \varepsilon/2$ for $n \geq N_1$ and
$|y_n - z_n| < \varepsilon/2$ for $n \geq N_2$.
For all $n \geq \max\{N_1, N_2\}$:
\[
|x_n - z_n|
\leq |x_n - y_n| + |y_n - z_n|
< \frac{\varepsilon}{2} + \frac{\varepsilon}{2} = \varepsilon.
\]
Therefore $(x_n) \sim (z_n)$.
\end{proof}

\begin{proof}[Proof (Detailed Learning)]
We verify the three axioms of an equivalence relation.

\textbf{Step 1: Reflexivity — $(x_n) \sim (x_n)$.}
We need: $\forall \varepsilon > 0$, $\exists N$ such that
$|x_n - x_n| < \varepsilon$ for all $n \geq N$.
Since $x_n - x_n = 0$ for every $n$, we have $|x_n - x_n| = 0 < \varepsilon$
for every $n$ and every $\varepsilon > 0$.  Any $N = 1$ works.

\textbf{Step 2: Symmetry — $(x_n) \sim (y_n)$ implies $(y_n) \sim (x_n)$.}
Suppose $(x_n) \sim (y_n)$.  Let $\varepsilon > 0$.  There exists $N$ such
that $|x_n - y_n| < \varepsilon$ for all $n \geq N$.
Since $|y_n - x_n| = |-(x_n - y_n)| = |x_n - y_n|$, we have
$|y_n - x_n| < \varepsilon$ for all $n \geq N$.
Therefore $(y_n) \sim (x_n)$.

\textbf{Step 3: Transitivity — $(x_n) \sim (y_n)$ and $(y_n) \sim (z_n)$
implies $(x_n) \sim (z_n)$.}
Suppose $(x_n) \sim (y_n)$ and $(y_n) \sim (z_n)$.
Let $\varepsilon > 0$.

\emph{Extract witnesses.}  Since $(x_n) \sim (y_n)$, there exists $N_1$
such that $|x_n - y_n| < \varepsilon/2$ for all $n \geq N_1$.
Since $(y_n) \sim (z_n)$, there exists $N_2$ such that
$|y_n - z_n| < \varepsilon/2$ for all $n \geq N_2$.

\emph{Combine.}  Set $N = \max\{N_1, N_2\}$.  For all $n \geq N$:
\[
|x_n - z_n|
= |x_n - y_n + y_n - z_n|
\leq |x_n - y_n| + |y_n - z_n|
< \frac{\varepsilon}{2} + \frac{\varepsilon}{2}
= \varepsilon.
\]

Therefore $(x_n) \sim (z_n)$.

\textbf{Step 4: Conclude.}
All three axioms hold, so $\sim$ is an equivalence relation on $\mathbb{Q}^c$.
\end{proof}

\begin{remark*}[Proof structure]
This proof is a direct three-part verification.  Reflexivity is trivial.
Symmetry follows from $|a - b| = |b - a|$.  Transitivity is a
$2\varepsilon$ split via the triangle inequality — structurally identical
to the proof of uniqueness of limits and the proof that convergent
sequences are Cauchy.  All three proofs share a single master template.
\end{remark*}

\begin{remark*}[Dependencies]
\hyperref[def:cauchy-equivalence]{Equivalence of Cauchy sequences},
triangle inequality in $\mathbb{Q}$.
\end{remark*}
