\newpage
\phantomsection
\label{prf:ec-integer-embedding-add}

\begin{remark*}[Return]
\hyperref[thm:ec-integer-embedding-add]{Return to Theorem}
\end{remark*}

\begin{theorem*}[Integer Embedding Preserves Addition]
For all $m,n\in\mathbb{Z}$,
\[
[(m,1)] + [(n,1)] = [(m+n,\,1)].
\]
\end{theorem*}

\begin{proof}
\textbf{Professional Standard Proof.}~\\
Let $m,n\in\mathbb{Z}$.  By the definition of addition on rational
equivalence classes,
\[
[(m,1)] + [(n,1)]
= [(m\cdot 1 + 1\cdot n,\; 1\cdot 1)]
= [(m+n,\; 1)].
\]
\end{proof}

\begin{proof}
\textbf{Detailed Learning Proof.}~\\

\textbf{Step 1.} Identify the embeddings.

Let $m,n\in\mathbb{Z}$.  Under the integer embedding, $m$ is identified
with $[(m,1)]\in\mathbb{Q}$ and $n$ is identified with $[(n,1)]\in\mathbb{Q}$.
Both pairs $(m,1)$ and $(n,1)$ are admissible fraction-pairs since $1\neq 0$.

\textbf{Step 2.} Apply the addition formula.

By the definition of addition on rational equivalence classes,
\[
[(m,1)] + [(n,1)]
:= [(m\cdot 1 + 1\cdot n,\; 1\cdot 1)].
\]

\textbf{Step 3.} Simplify using integer arithmetic.

In $\mathbb{Z}$:
\[
m\cdot 1 + 1\cdot n = m + n
\qquad\text{and}\qquad
1\cdot 1 = 1.
\]
Therefore
\[
[(m\cdot 1 + 1\cdot n,\; 1\cdot 1)] = [(m+n,\; 1)].
\]

\textbf{Step 4.} Conclude.

Combining Steps 2 and 3:
\[
[(m,1)] + [(n,1)] = [(m+n,\,1)].
\]
This is the embedding of the integer $m+n$, so the embedding preserves
addition.
\end{proof}

\begin{remark*}[Proof structure]
The proof is a single application of the addition formula from the definition
of rational operations, followed by integer arithmetic.  There is no
equivalence relation work and no well-definedness check --- those were
discharged once and for all by the well-definedness lemma when the
operations were defined.  The sole content is reading off the formula at
the specific pair $(m,1)$, $(n,1)$ and simplifying $m\cdot 1 + 1\cdot n$
to $m+n$.
\end{remark*}

\begin{remark*}[Dependencies]~\\
\begin{itemize}
  \item \hyperref[def:rational-operations]{Definition~\ref*{def:rational-operations}}
        \textnormal{(addition formula: $[(a,b)]+[(c,d)]:=[(ad+bc,bd)]$)}
  \item \hyperref[lem:rational-addition-well-defined]{Lemma~\ref*{lem:rational-addition-well-defined}}
        \textnormal{(well-definedness of addition, discharges representative independence)}
  \item Basic integer arithmetic ($m\cdot 1 = m$, $1\cdot n = n$, $1\cdot 1 = 1$)
\end{itemize}
\end{remark*}
