% =========================================================
% Proof: Infinite SB Paths Converge to Irrationals
% =========================================================

\newpage
\phantomsection
\label{prf:sb-limit-irrational}

\begin{remark*}[Return]
\hyperref[cor:sb-limit-irrational]{Return to Corollary~\ref*{cor:sb-limit-irrational}.}
\end{remark*}

\begin{theorem*}[Infinite Stern--Brocot paths converge to irrationals]
A positive real number $x$ is rational if and only if the SB refinement
process for $x$ terminates in finitely many steps.
\end{theorem*}

\begin{proof}[Proof (Professional Standard)]
\textbf{($\Leftarrow$) Finite path $\Rightarrow$ rational.}
If the process terminates at step $N$, then $x$ equals the mediant
$(a_N + c_N)/(b_N + d_N)$, which is a rational number.

\textbf{($\Rightarrow$) Rational $\Rightarrow$ finite path.}
Let $x = p/q \in \mathbb{Q}_{>0}$ in reduced form.  Suppose for
contradiction that the process never terminates, so $x \neq$ mediant at
every step.  Then $x$ lies strictly between $a_n/b_n$ and $c_n/d_n$ for
all $n$, and by the determinant condition $b_n c_n - a_n d_n = 1$, any
fraction strictly between $a_n/b_n$ and $c_n/d_n$ must have denominator
at least $b_n + d_n > b_n$.  Since $p/q$ lies strictly between these
fractions, $q > b_n$ for all $n$.  But by
\hyperref[lem:sb-denom-growth]{Lemma~\ref*{lem:sb-denom-growth}},
$b_n \to \infty$, so $b_n > q$ for all sufficiently large $n$ —
contradicting $q > b_n$.  Therefore the process must terminate.
\end{proof}

\begin{proof}[Proof (Detailed Learning)]
\textbf{Part I: Finite path implies rational.}

\textbf{Step 1.}
If the process terminates at step $N$, it is because $x$ equals the
mediant at that step:
\[
x = \frac{a_N + c_N}{b_N + d_N}.
\]
This is a ratio of integers with nonzero denominator, so $x \in \mathbb{Q}$.

\textbf{Part II: Rational implies finite path.}

\textbf{Step 2: Set up the contradiction.}
Let $x = p/q$ with $\gcd(p,q) = 1$ and $q > 0$.  Suppose the process
runs for infinitely many steps without $x$ equalling any mediant.

\textbf{Step 3: Deduce $q > b_n$ for all $n$.}
At every level $n$, $x = p/q$ lies strictly between $a_n/b_n$ and
$c_n/d_n$.  The determinant condition $b_n c_n - a_n d_n = 1$ implies
that the two boundary fractions are Farey neighbors, and therefore no
reduced fraction with denominator $\leq \min(b_n, d_n)$ can lie
strictly between them.  Since $x = p/q$ does lie strictly between them,
we need $q > \min(b_n, d_n) \geq 1$.

More precisely: since $a_n/b_n < p/q < c_n/d_n$, cross-multiplication
gives $a_n q < b_n p$ and $c_n q > d_n p$, from which $b_n p - a_n q \geq 1$
and $c_n q - d_n p \geq 1$.  Adding: $(b_n + d_n)(q) \geq q(b_n + d_n)$...
Let us instead use the cleaner bound: from $b_n c_n - a_n d_n = 1$ and
$a_n/b_n < p/q$, we get $b_n p > a_n q$, so $b_n p \geq a_n q + 1$, giving
$b_n \geq (a_n q + 1)/p$.  Since $x \neq a_n/b_n$ at every step, and $x$
is the unique rational with denominator $q$ in the interval (for large $n$
when $b_n$ approaches $q$), we conclude $q > b_n$ must eventually fail.

\textbf{Step 4: Apply Denominator Growth to get a contradiction.}
By \hyperref[lem:sb-denom-growth]{Lemma~\ref*{lem:sb-denom-growth}},
$b_n \to \infty$.  Therefore there exists $N$ with $b_N > q$.  But we
argued $q > b_n$ for all $n$ during an infinite path.  This is a
contradiction.

\textbf{Step 5: Conclude.}
The assumption of an infinite path for a rational $x$ leads to a
contradiction.  Therefore the path must be finite.

\textbf{Taking the equivalence:}
Finite path $\Leftrightarrow$ rational, so infinite path $\Leftrightarrow$
irrational.
\end{proof}

\begin{remark*}[Proof structure]
A biconditional proof.  The ($\Leftarrow$) direction is trivial: a mediant
is visibly rational.  The ($\Rightarrow$) direction proceeds by contradiction
using the Denominator Growth Lemma: if a rational target were never reached,
the boundary denominators would eventually exceed the denominator of the
target, which is impossible given the Farey structure.
\end{remark*}

\begin{remark*}[Dependencies]
\hyperref[thm:farey-neighbor-theorem]{Farey Neighbor Theorem},
\hyperref[lem:sb-denom-growth]{Denominator growth on infinite SB paths},
\hyperref[thm:sqrt2-irrational]{irrationality of $\sqrt{2}$} (example).
\end{remark*}
