\newpage
\phantomsection
\label{prf:power-of-product-in-field}

\begin{remark*}[Return]
\hyperref[thm:power-of-product-in-field]{Return to Theorem}
\end{remark*}

\begin{theorem*}[Power of a Product in a Field]
Let $(F,+,\cdot,0,1)$ be a field. Let $q,r\in F\setminus\{0\}$ and let
$a\in\mathbb{Z}$. Then
\[
(qr)^a=q^a r^a.
\]
\end{theorem*}

\begin{proof}
\textbf{Professional Standard Proof.}~\\
First suppose that $a\ge 0$. Then $a\in\mathbb{N}$, and the usual
law of exponents for natural-number powers gives
\[
(qr)^a=q^a r^a.
\]

Now suppose that $a<0$. Then there exists $n\in\mathbb{N}$ with
$n>0$ such that
\[
a=-n.
\]
Since $q\neq 0$ and $r\neq 0$, their product is nonzero:
\[
qr\neq 0.
\]
By the definition of negative integer powers,
\[
(qr)^a=(qr)^{-n}=\frac{1}{(qr)^n}.
\]
Using the natural-number power law from the first case,
\[
(qr)^n=q^n r^n.
\]
Hence
\[
(qr)^a=\frac{1}{q^n r^n}.
\]
Since $q^n\neq 0$ and $r^n\neq 0$, the inverse of the product is the
product of the inverses:
\[
\frac{1}{q^n r^n}=\frac{1}{q^n}\frac{1}{r^n}.
\]
By the definition of negative integer powers,
\[
\frac{1}{q^n}=q^{-n}
\qquad\text{and}\qquad
\frac{1}{r^n}=r^{-n}.
\]
Therefore
\[
(qr)^a=q^{-n}r^{-n}=q^a r^a.
\]
Thus, for every $a\in\mathbb{Z}$,
\[
(qr)^a=q^a r^a.
\]
\end{proof}

\begin{proof}
\textbf{Detailed Learning Proof.}~\\

\textbf{Step 1.} Prove the nonnegative exponent case.

Assume first that $a\ge 0$. Then $a\in\mathbb{N}$. For natural-number
powers, the product law for exponents gives
\[
(qr)^a=q^a r^a.
\]
This proves the result for all nonnegative integer exponents.

\textbf{Step 2.} Reduce the negative exponent case to a positive exponent.

Assume now that $a<0$. Then there exists $n\in\mathbb{N}$ with
$n>0$ such that
\[
a=-n.
\]
Since $q,r\in F\setminus\{0\}$, both $q$ and $r$ are nonzero. Because
$F$ is a field, the product of two nonzero elements is nonzero, so
\[
qr\neq 0.
\]
Therefore the negative power $(qr)^{-n}$ is defined.

\textbf{Step 3.} Expand the negative power of the product.

By the definition of negative integer exponents,
\[
(qr)^a=(qr)^{-n}=\frac{1}{(qr)^n}.
\]
Since $n\in\mathbb{N}$, the natural-number product law gives
\[
(qr)^n=q^n r^n.
\]
Substituting this into the previous equality gives
\[
(qr)^a=\frac{1}{q^n r^n}.
\]

\textbf{Step 4.} Split the inverse of the product.

Because $q\neq 0$ and $r\neq 0$, the powers $q^n$ and $r^n$ are
nonzero. Hence
\[
q^n r^n\neq 0.
\]
In a field, the inverse of a product of nonzero elements is the product
of the inverses. Thus
\[
\frac{1}{q^n r^n}=\frac{1}{q^n}\frac{1}{r^n}.
\]

\textbf{Step 5.} Convert the inverses back into negative powers.

By the definition of negative integer exponents,
\[
\frac{1}{q^n}=q^{-n}
\qquad\text{and}\qquad
\frac{1}{r^n}=r^{-n}.
\]
Therefore
\[
(qr)^a
=
\frac{1}{q^n r^n}
=
\frac{1}{q^n}\frac{1}{r^n}
=
q^{-n}r^{-n}.
\]
Since $a=-n$, this becomes
\[
(qr)^a=q^a r^a.
\]
This proves the theorem.
\end{proof}

\begin{remark*}[Proof structure]
The proof proceeds by cases on the integer exponent. For nonnegative
exponents, the result is the usual natural-number power law. For negative
exponents, the exponent is written as $-n$ with $n>0$, the negative
power is converted into an inverse of a positive power, the natural-number
case is applied, and then the inverse of a product is split into the product
of inverses.
\end{remark*}

\begin{remark*}[Dependencies]~\\
\begin{itemize}
  \item \hyperref[def:negative-integer-exponents]{Definition~\ref*{def:negative-integer-exponents}}
  \item \hyperref[def:field]{Definition~\ref*{def:field}}
  \item Natural-number product law for powers
  \item Nonzero product property in a field
  \item Inverse of a product in a field
\end{itemize}
\end{remark*}