% =========================================================
% Proof: Stern--Brocot Enumeration
% =========================================================

\newpage
\phantomsection
\label{prf:stern-brocot-enumeration}

\begin{remark*}[Return]
\hyperref[thm:stern-brocot-enumeration]{Return to Theorem~\ref*{thm:stern-brocot-enumeration}.}
\end{remark*}

\begin{theorem*}[Stern--Brocot enumeration]
Every positive rational number appears exactly once in reduced form in
the Stern--Brocot tree.
\end{theorem*}

\begin{proof}[Proof (Professional Standard)]
We prove existence and uniqueness separately.

\textbf{Existence.} Let $p/q \in \mathbb{Q}_{>0}$ be in reduced form.
The Stern--Brocot search process starts from $0/1$ and $1/0$ and at each
step compares $p/q$ to the current mediant $m$.  If $p/q < m$, go left;
if $p/q > m$, go right; if $p/q = m$, terminate.  At each step, the two
boundary fractions $a/b$ and $c/d$ satisfy $bc - ad = 1$
(\hyperref[thm:farey-neighbor-theorem]{Farey Neighbor Theorem}).  The
mediant has denominator $b + d$.  Since $p/q$ lies strictly between $a/b$
and $c/d$ at every non-terminal step, a descent argument (analogous to
the Euclidean algorithm on the pair $(b, d)$) shows the denominators of
the successive boundaries strictly decrease toward $q$, forcing the process
to terminate at $p/q$ in finitely many steps.
Therefore every positive rational is reached.

\textbf{Uniqueness.} Suppose $p/q$ appears at two distinct nodes.  Each
node is reached by a unique left-right path from the root, and the
Stern--Brocot structure guarantees that the fractions encountered along
any path are strictly increasing (when going right) or strictly decreasing
(when going left).  Two distinct paths reaching $p/q$ would require the
search to visit $p/q$ twice with the same value but different surrounding
intervals, which is impossible given the determinant condition
$bc - ad = 1$ that is maintained at every level.  Therefore every positive
rational appears exactly once.
\end{proof}

\begin{proof}[Proof (Detailed Learning)]
\textbf{Step 1: Describe the search procedure.}
Start with the boundary pair $(a/b, c/d) = (0/1, 1/0)$.
At each step compute the mediant $m = (a+c)/(b+d)$ and compare $p/q$ to $m$:
if $p/q = m$, terminate; if $p/q < m$, set $(a/b, c/d) \leftarrow (a/b, m)$;
if $p/q > m$, set $(a/b, c/d) \leftarrow (m, c/d)$.

\textbf{Step 2: The determinant condition is preserved.}
Initially $bc - ad = 1 \cdot 1 - 0 \cdot 0 = 1$.  When we replace
$c/d$ by the mediant $(a+c)/(b+d)$, the new determinant is
$b(a+c) - a(b+d) = bc - ad = 1$.  Similarly for replacing $a/b$.
Therefore $bc - ad = 1$ holds at every level.

\textbf{Step 3: Existence via denominator descent.}
At each step where $p/q \neq m$, we have $p/q$ strictly between $a/b$
and $c/d$, so $bq - ap \geq 1$ and $cp - dq \geq 1$.
The new boundary after the step has a smaller denominator on the side
we replaced (since the mediant's denominator is $b + d$, but the
replaced boundary had denominator $\leq q - 1$ when $p/q = p/q$).
By a descent argument on $q - b$ and $q - d$ (which are positive integers
that strictly decrease at each step), the process must terminate in at
most $2q - 1$ steps.

\textbf{Step 4: Uniqueness.}
Any node in the tree is the unique mediant of its two parent boundaries
at the preceding level.  Since $bc - ad = 1$ pins the arithmetic
relationship between successive boundaries uniquely, and the search
direction at each step is determined by the comparison $p/q$ vs.\ mediant,
there is only one path from the root to any given value.  Two distinct
paths to the same rational would require two different sequences of
left-right decisions, but the comparison is deterministic, so this is
impossible.

\textbf{Step 5: Conclude.}
Every positive rational appears (existence) and appears exactly once
(uniqueness).
\end{proof}

\begin{remark*}[Proof structure]
Existence uses a descent argument analogous to the Euclidean algorithm.
Uniqueness uses the deterministic nature of the search.  The Farey
determinant condition $bc - ad = 1$ is the structural invariant that
makes both arguments work.
\end{remark*}

\begin{remark*}[Dependencies]
\hyperref[def:mediant]{Mediant},
\hyperref[thm:farey-neighbor-theorem]{Farey Neighbor Theorem},
\hyperref[lem:mediant-gap]{Mediant Gap Lemma},
\hyperref[thm:fraction-pair-equivalence]{reduced-form representatives}.
\end{remark*}
