% =========================================================
% Breadcrumb
% =========================================================
\begin{tcolorbox}[
  colback=gray!6,
  colframe=gray!40,
  arc=2pt,
  left=8pt, right=8pt, top=6pt, bottom=6pt,
  title={\small\textbf{Where You Are in the Journey}},
  fonttitle=\small\bfseries
]
\begin{center}
\small
Propositional Logic
$\;\to\;$ Predicate Calculus
$\;\to\;$ Sets \& Functions
$\;\to\;$ Proof Techniques
$\;\to\;$ Axiom Systems
$\;\to\;$ $\mathbb{N}$, $\mathbb{Z}$, $\mathbb{Q}$
$\;\to\;$ \textbf{Real Numbers ($\mathbb{R}$)}
$\;\to\;$ Real Analysis
$\;\to\;$ $\cdots$
\end{center}

\medskip
\noindent\textbf{How we got here.}
The rational numbers gave us a dense ordered field, but they contain gaps:
there is no $q \in \mathbb{Q}$ with $q^2 = 2$.
The real number system fills these gaps. Two rigorous constructions achieve
this --- Dedekind cuts and Cauchy sequences --- and both yield the same
unique complete ordered field.

\medskip
\noindent\textbf{What this chapter builds.}
The axiomatic structure of $\mathbb{R}$: field axioms, order axioms,
the completeness axiom, and their immediate consequences ---
the Archimedean property, density of $\mathbb{Q}$, and the
interval and bound theory that underpins all of analysis.
We also study how $\mathbb{R}$ is \emph{constructed} from $\mathbb{Q}$
via Dedekind cuts and Cauchy sequences.

\medskip
\noindent\textbf{Where this leads.}
Real analysis in Volume~III takes $\mathbb{R}$ as given and develops
sequences, series, convergence, and their limiting behaviour.
The field and order axioms established here propagate unchanged
into metric spaces, normed spaces, and abstract algebra.
\end{tcolorbox}
\vspace{1em}

% =========================================================
% Structural Roadmap
% =========================================================
\section*{Structural Roadmap}

This chapter covers the \emph{algebraic and analytic axioms} of
$\mathbb{R}$ and the \emph{constructions} that justify its existence.
Sequence and series theory is developed in
\textbf{Volume~III, Real Analysis}.

\begin{center}
\textbf{Axioms $\longrightarrow$ Order \& Bounds
$\longrightarrow$ Completeness $\longrightarrow$ Constructions}
\end{center}

The global progression is:
\begin{enumerate}
  \item Field and order axioms
  \item Intervals and convexity
  \item Bounds: upper/lower bounds, maximum, minimum
  \item Extremal values: supremum and infimum
  \item Completeness: the least upper bound property and its consequences
  \item Constructions of $\mathbb{R}$: Dedekind cuts
  \item Constructions of $\mathbb{R}$: Cauchy sequences
  \item Interval arithmetic
\end{enumerate}

\vspace{1em}

\begin{remark}[Primary sources]
Axiomatic development follows Abbott, \textit{Understanding Analysis}
and Ross, \textit{Elementary Analysis}.
The Cauchy sequence construction follows Tao, \textit{Analysis~I},
Chapters~5--6.
The Dedekind cut construction follows Rudin,
\textit{Principles of Mathematical Analysis}, Appendix to Chapter~1.
\end{remark}

% =========================================================
% Content
% =========================================================
\input{volume-ii/reals/notes/index}
