\newpage
\phantomsection
\label{prf:fraction-pair-equivalence}
\LRAProofFor{thm:fraction-pair-equivalence}

\begin{remark*}[Return]
\hyperref[thm:fraction-pair-equivalence]{Return to Theorem}
\end{remark*}

\begin{theorem*}[The Fraction-Pair Relation Is an Equivalence Relation]
The relation $\sim$ on $W$ defined by
\[
(a,b)\sim(c,d)\iff ad=bc
\]
is an equivalence relation.
\end{theorem*}

\begin{proof}[Professional Standard Proof]
\LRAProofBodyStart
Assume
\[
W=\mathbb{Z}\times(\mathbb{Z}\setminus\{0\}).
\]
Thus every element of $W$ has the form $(a,b)$ with $a,b\in\mathbb{Z}$
and $b\neq 0$.

First, let $(a,b)\in W$. Then
\[
ab=ba
\]
by commutativity of integer multiplication. Hence
\[
(a,b)\sim(a,b).
\]
Therefore $\sim$ is reflexive.

Next, let $(a,b),(c,d)\in W$ and suppose
\[
(a,b)\sim(c,d).
\]
By definition,
\[
ad=bc.
\]
Therefore
\[
cb=da
\]
by symmetry of equality and commutativity of integer multiplication. Hence
\[
(c,d)\sim(a,b).
\]
Therefore $\sim$ is symmetric.

Finally, let $(a,b),(c,d),(e,f)\in W$ and suppose
\[
(a,b)\sim(c,d)
\qquad\text{and}\qquad
(c,d)\sim(e,f).
\]
By definition,
\[
ad=bc
\qquad\text{and}\qquad
cf=de.
\]
Multiplying the first equality by $f$ gives
\[
adf=bcf.
\]
Multiplying the second equality by $b$ gives
\[
bcf=bde.
\]
Hence
\[
adf=bde.
\]
By associativity and commutativity of integer multiplication,
\[
d(af)=d(be).
\]
Since $(c,d)\in W$, we have $d\neq 0$. By cancellation in the integers,
\[
af=be.
\]
Therefore
\[
(a,b)\sim(e,f).
\]
Thus $\sim$ is transitive.

Since $\sim$ is reflexive, symmetric, and transitive, it is an equivalence
relation on $W$.
\end{proof}

\begin{proof}[Detailed Learning Proof]
\LRAProofBodyStart

\textbf{Step 1.} Verify reflexivity.

Let $(a,b)\in W$. Since $W=\mathbb{Z}\times(\mathbb{Z}\setminus\{0\})$,
we have $a,b\in\mathbb{Z}$ and $b\neq 0$. To prove reflexivity, it is
necessary to show
\[
(a,b)\sim(a,b).
\]
By the definition of $\sim$, this is equivalent to showing
\[
ab=ba.
\]
This equality holds by commutativity of multiplication in $\mathbb{Z}$.
Therefore
\[
(a,b)\sim(a,b).
\]
Hence $\sim$ is reflexive.

\textbf{Step 2.} Verify symmetry.

Let $(a,b),(c,d)\in W$, and assume
\[
(a,b)\sim(c,d).
\]
By the definition of $\sim$, this means
\[
ad=bc.
\]
To prove symmetry, it is necessary to show
\[
(c,d)\sim(a,b).
\]
By the definition of $\sim$, this is equivalent to showing
\[
cb=da.
\]
From $ad=bc$, symmetry of equality gives
\[
bc=ad.
\]
By commutativity of multiplication in $\mathbb{Z}$,
\[
cb=bc
\qquad\text{and}\qquad
ad=da.
\]
Therefore
\[
cb=da.
\]
Hence
\[
(c,d)\sim(a,b).
\]
Thus $\sim$ is symmetric.

\textbf{Step 3.} Verify transitivity.

Let $(a,b),(c,d),(e,f)\in W$, and assume
\[
(a,b)\sim(c,d)
\qquad\text{and}\qquad
(c,d)\sim(e,f).
\]
By the definition of $\sim$, these assumptions give
\[
ad=bc
\qquad\text{and}\qquad
cf=de.
\]
To prove transitivity, it is necessary to show
\[
(a,b)\sim(e,f),
\]
which means
\[
af=be.
\]

Multiplying the equality $ad=bc$ by $f$ gives
\[
adf=bcf.
\]
Multiplying the equality $cf=de$ by $b$ gives
\[
bcf=bde.
\]
Combining these two equalities gives
\[
adf=bde.
\]
Using associativity and commutativity of multiplication in $\mathbb{Z}$,
this may be rewritten as
\[
d(af)=d(be).
\]
Since $(c,d)\in W$, the second coordinate $d$ is nonzero. Therefore
cancellation by the nonzero integer $d$ gives
\[
af=be.
\]
Hence
\[
(a,b)\sim(e,f).
\]
Thus $\sim$ is transitive.

\textbf{Step 4.} Conclude that the relation is an equivalence relation.

The relation $\sim$ has been shown to be reflexive, symmetric, and
transitive. Therefore $\sim$ is an equivalence relation on $W$.
\end{proof}

\begin{remark*}[Proof structure]
The proof is direct. Reflexivity follows from commutativity of integer
multiplication. Symmetry follows by reversing the equality and commuting
the products. Transitivity is the only substantive step: two cross-product
equalities are multiplied by the remaining denominators, combined, and
then cancellation by a nonzero integer produces the required cross-product
equality.
\end{remark*}

\begin{dependencies}
\begin{itemize}
  \item \hyperref[def:fraction-pair-equality]{Definition~\ref*{def:fraction-pair-equality}}
  \item Basic arithmetic of the integers
  \item Cancellation law for multiplication by a nonzero integer
\end{itemize}
\end{dependencies}
