\newpage
\phantomsection
\label{prf:multiply-positive}
\LRAProofFor{thm:multiply-positive}

\begin{remark*}[Return]
\hyperref[thm:multiply-positive]{Return to Theorem}
\end{remark*}

\begin{theorem*}[Multiplication by a Positive Element Preserves Order]
If $a<b$ and $c>0$, then
\[
ac<bc.
\]
\end{theorem*}

\begin{proof}
\textbf{Professional Standard Proof.}~\\
Let $F$ be an ordered field, and suppose that
\[
a<b
\qquad\text{and}\qquad
c>0.
\]
By definition, an ordered field is a field equipped with an order making it
an ordered integral domain.  One of the axioms of an ordered integral domain
states precisely that whenever
\[
a<b
\qquad\text{and}\qquad
0<c,
\]
then
\[
ac<bc.
\]
Applying this axiom to the present hypotheses gives
\[
ac<bc.
\]
This is exactly the desired conclusion.
\end{proof}

\begin{remark*}[Proof structure]
This proof is an unpacking of definitions.  The theorem is not derived from
earlier algebraic lemmas; it is one of the compatibility clauses built into
the definition of an ordered integral domain, and hence into the definition
of an ordered field.  So the argument consists of recognizing that the
hypotheses match the axiom exactly and then invoking it.
\end{remark*}

\begin{dependencies}
\begin{itemize}
  \item \hyperref[def:ordered-field]{Definition~\ref*{def:ordered-field}}
  \item \hyperref[def:ordered-integral-domain]{Definition~\ref*{def:ordered-integral-domain}}
\end{itemize}
\end{dependencies}

\textbf{Detailed Learning Proof.}~\\
% MIGRATION TODO: Existing proof content preserved; detailed/professional companion body needs review.
TODO: Expand the proof into a detailed learning proof.

