% =========================================================
% Archived live roadmap: previous natural-numbers chapter
% =========================================================
\begin{tcolorbox}[
  colback=gray!6,
  colframe=gray!40,
  arc=2pt,
  left=8pt, right=8pt, top=6pt, bottom=6pt,
  title={\small\textbf{Where You Are in the Journey}},
  fonttitle=\small\bfseries
]
\begin{center}
\small
Propositional Logic
$\;\to\;$ Sets \& Functions
$\;\to\;$ Proof Techniques
$\;\to\;$ Axiom Systems
$\;\to\;$ \textbf{Natural Numbers ($\mathbb{N}$)}
$\;\to\;$ Integers ($\mathbb{Z}$)
$\;\to\;$ Rationals ($\mathbb{Q}$)
$\;\to\;$ Reals ($\mathbb{R}$)
$\;\to\;$ $\cdots$
\end{center}

\medskip
\noindent\textbf{How we got here.}
The Peano axioms in Volume~I established that $\mathbb{N}$ exists:
five axioms pin down the structure of zero and the successor function,
and the recursion theorem guarantees that recursive definitions
actually produce well-defined functions.
This chapter takes those logical primitives and builds arithmetic ---
the concrete mathematics of counting, addition, multiplication, and order.

\medskip
\noindent\textbf{What this chapter builds.}
We follow Tao's \emph{Analysis~I}, Chapter~2 (§§2.2--2.3).
Addition is defined by recursion on the left argument and every
algebraic property (commutativity, associativity, cancellation)
is proved from the definition alone.
Order on $\mathbb{N}$ is defined additively and the full trichotomy
is established. Multiplication is then defined recursively
and its properties derived. Strong induction and the
well-ordering principle round out the toolkit.

\medskip
\noindent\textbf{Where this leads.}
The integers $\mathbb{Z}$ are constructed by adjoining additive inverses
to $\mathbb{N}$.
The cancellation law for natural-number addition --- proved here ---
carries the entire integer construction forward.
Every algebraic structure built later inherits the patterns
established in this chapter.
\end{tcolorbox}
\vspace{1em}

\section*{Structural Roadmap}

The development follows Tao's construction of $\mathbb{N}$ arithmetic
from the Peano axioms. The architecture at every step is:
\begin{center}
\textbf{Definition (recursive) $\longrightarrow$ Well-definedness
$\longrightarrow$ Lemmas $\longrightarrow$ Propositions
$\longrightarrow$ Corollaries}
\end{center}

Nothing beyond the Peano axioms and what has already been proved
is assumed. The available toolkit grows with each result.

The global progression is:
\begin{enumerate}
  \item \textbf{Addition} (Tao §2.2, Def.~2.2.1):
        defined by recursion on the left argument.
        Key results: zero on the right, successor on the right,
        commutativity, associativity, cancellation.
  \item \textbf{Order on $\mathbb{N}$} (Tao §2.2, Def.~2.2.11):
        $n \leq m$ iff $\exists a \in \mathbb{N},\ m = n + a$.
        Key results: transitivity, antisymmetry, trichotomy,
        monotonicity under addition.
  \item \textbf{Strong induction} (Tao P2.2.14):
        the principle of strong induction and backwards induction,
        derived from the Peano induction axiom.
  \item \textbf{Multiplication} (Tao §2.3, Def.~2.3.1):
        defined by recursion. Key results: commutativity,
        associativity, distributivity, no zero divisors.
  \item \textbf{Exponentiation} (Mendelson §2.6):
        defined recursively. Basic exponent laws proved.
\end{enumerate}

\vspace{1em}

\begin{remark}[What stays in Axiom Systems]
The Peano axioms (P1--P5), numeral definitions ($1 := 0\pp$, etc.),
and the recursion theorem live in Volume~I, Axiom Systems.
That chapter is the \emph{logical} foundation; this chapter is the
\emph{arithmetic} development built on top of it.
\end{remark}

\begin{remark}[Primary sources]
Tao, \textit{Analysis~I}, Chapter~2 (primary);
Mendelson, \textit{Number Systems and the Foundations of Analysis},
Chapter~2 (exponentiation and additional structural perspective).
\end{remark}
