% =========================================================
% Gaps in the Rational Numbers
% =========================================================

\subsection{Gaps in the Rational Numbers}
\label{sec:gaps-in-q}

The rational numbers are dense and Archimedean, yet they are not complete.
This means that there are real-number locations that can be approached
arbitrarily closely by rationals but are not occupied by any rational number.
The most famous example is $\sqrt{2}$.

% ---------------------------------------------------------
% Theorem: No Rational Number Has Square 2
% ---------------------------------------------------------
\begin{tcolorbox}[colback=thmbox, colframe=thmborder, arc=2pt,
  left=6pt, right=6pt, top=4pt, bottom=4pt,
  title={\small\textbf{Theorem (No Rational Number Has Square $2$)}},
  fonttitle=\small\bfseries]
\begin{theorem}[No Rational Number Has Square $2$]
\label{thm:sqrt2-irrational}
There is no rational number whose square is $2$.
\hyperref[prf:sqrt2-irrational]{Go to proof.}
\end{theorem}
\end{tcolorbox}

\begin{remark*}[Standard quantified statement]
\[
\neg\exists r\in\mathbb{Q}\;\bigl(r^2=2\bigr).
\]
\end{remark*}

\begin{remark*}[Predicate reading]
\[
\forall r\in\mathbb{Q}\;\neg\bigl(r^2=2\bigr).
\]
\end{remark*}

\begin{remark*}[Negated quantified statement]
\[
\exists r\in\mathbb{Q}\;\bigl(r^2=2\bigr).
\]
\end{remark*}

\begin{remark*}[Failure modes]
The theorem would fail exactly if some rational number were an exact square
root of $2$. Writing it in lowest terms as $p/q$ with $\gcd(p,q)=1$ leads to
$p^2=2q^2$, forcing $p$ even and then $q$ even --- contradicting $\gcd(p,q)=1$.
\end{remark*}

\begin{remark*}[Failure mode decomposition]
\[
\neg\bigl(\forall r\in\mathbb{Q}\;\neg(r^2=2)\bigr)
\Longleftrightarrow
\exists r\in\mathbb{Q}\;(r^2=2).
\]
\end{remark*}

\begin{remark*}[Interpretation]
This is the first visible gap in $\mathbb{Q}$. Rational numbers can
approximate $\sqrt{2}$ arbitrarily well, but no rational number lands exactly
on it. The obstruction is arithmetic: squaring a reduced rational cannot
produce $2$ without violating the parity structure of the integers.
\end{remark*}

\begin{remark*}[Dependencies]
\begin{itemize}
  \item \hyperref[def:rational-numbers]{Rational Numbers}
  \item \hyperref[thm:fraction-pair-equivalence]{Fraction Pair Equivalence}
\end{itemize}
\end{remark*}

\begin{remark}[The gap at $\sqrt{2}$]
Consider the sets
\[
A = \{r\in\mathbb{Q}: r^2<2\}
\qquad\text{and}\qquad
B = \{r\in\mathbb{Q}: r^2>2\}.
\]
Every rational number lies in one of these sets. Numbers in $A$ approximate
$\sqrt{2}$ from below and numbers in $B$ approximate it from above. But there
is no rational number that fills the gap between them.
\end{remark}

% ---------------------------------------------------------
% Lemma: Rational Square Bracketing
% ---------------------------------------------------------
\begin{lemma}[Rational Square Bracketing]
\label{lem:rational-square-bracketing}
Let $r$ be a positive rational number. Then there exist positive rational
numbers $a$ and $b$ such that
\[
a^2<r<b^2.
\]
\hyperref[prf:rational-square-bracketing]{Go to proof.}
\end{lemma}

\begin{remark*}[Standard quantified statement]
\[
\forall r\in\mathbb{Q},\;r>0\;\;
\exists a,b\in\mathbb{Q},\;a>0,\;b>0\;\;
a^2 < r < b^2.
\]
\end{remark*}

\begin{remark*}[Interpretation]
Every positive rational lies strictly between the squares of two positive
rationals. The witnesses are $a = r/(r+1)$ and $b = r+1$.
\end{remark*}

\begin{remark*}[Dependencies]
\begin{itemize}
  \item \hyperref[thm:q-is-ordered-field]{$\mathbb{Q}$ Is an Ordered Field}
\end{itemize}
\end{remark*}

% ---------------------------------------------------------
% Theorem: Rational Square Approximation from Above
% ---------------------------------------------------------
\begin{tcolorbox}[colback=thmbox, colframe=thmborder, arc=2pt,
  left=6pt, right=6pt, top=4pt, bottom=4pt,
  title={\small\textbf{Theorem (Rational Square Approximation from Above)}},
  fonttitle=\small\bfseries]
\begin{theorem}[Rational Square Approximation from Above]
\label{thm:rational-square-approximation-from-above}
Let $r$ be a positive rational number and let $n\in\mathbb{N}$. Then there
exists a positive rational number $s$ such that
\[
r<s^2<r+\frac{1}{n}.
\]
\hyperref[prf:rational-square-approximation-from-above]{Go to proof.}
\end{theorem}
\end{tcolorbox}

\begin{remark*}[Standard quantified statement]
\[
\forall r\in\mathbb{Q},\;r>0\;\;
\forall n\in\mathbb{N}\;\;
\exists s\in\mathbb{Q},\;s>0\;\;
r < s^2 < r+\tfrac{1}{n}.
\]
\end{remark*}

\begin{remark*}[Interpretation]
Positive rational numbers can be approximated from above by rational squares
with arbitrarily small excess. Together with the analogous result from below,
this means $r$ is the limit of a sequence of rational squares on both sides,
yet $r$ itself may not be a rational square.
\end{remark*}

\begin{remark*}[Dependencies]
\begin{itemize}
  \item \hyperref[lem:rational-square-bracketing]{Rational Square Bracketing}
  \item \hyperref[thm:archimedean-q]{Archimedean Property of $\mathbb{Q}$}
  \item \hyperref[thm:well-ordering-principle]{Well-Ordering Principle}
\end{itemize}
\end{remark*}

\begin{remark}
This is the decisive failure of $\mathbb{Q}$ as a number system for analysis.
The rational numbers support arithmetic, order, density, mediants, and powerful
approximation theorems. Even so, they still miss certain limits. The real
numbers are introduced precisely to repair this defect.
\end{remark}
