\newpage
\phantomsection
\label{prf:ec-scaling}
\LRAProofFor{thm:ec-scaling}

\begin{remark*}[Return]
\hyperref[thm:ec-scaling]{Return to Theorem}
\end{remark*}

\begin{theorem*}[Scaling of Equivalence Classes]
For every $[(a,b)]\in\mathbb{Q}$ and every nonzero integer $e\neq 0$,
\[
[(a,b)] = [(ae,be)].
\]
\end{theorem*}

\begin{proof}[Professional Standard Proof]
\LRAProofBodyStart
Let $[(a,b)]\in\mathbb{Q}$, and let $e\in\mathbb{Z}$ with $e\neq 0$.
Then $a,b\in\mathbb{Z}$ and $b\neq 0$. Since $b\neq 0$ and $e\neq 0$,
we also have
\[
be\neq 0.
\]
Thus $(ae,be)$ is an admissible fraction-pair.

It remains to show that
\[
(a,b)\sim(ae,be).
\]
By the definition of the fraction-pair relation, this is equivalent to showing
\[
a(be)=b(ae).
\]
Using associativity and commutativity of integer multiplication,
\[
a(be)=(ab)e=(ba)e=b(ae).
\]
Therefore
\[
(a,b)\sim(ae,be).
\]
Hence the corresponding equivalence classes are equal:
\[
[(a,b)] = [(ae,be)].
\]
\end{proof}

\begin{proof}[Detailed Learning Proof]
\LRAProofBodyStart

\textbf{Step 1.} Check that the scaled pair is admissible.

Let $[(a,b)]\in\mathbb{Q}$, and let $e\in\mathbb{Z}$ with
\[
e\neq 0.
\]
Since $[(a,b)]\in\mathbb{Q}$, the pair $(a,b)$ is an admissible
fraction-pair. Hence
\[
a,b\in\mathbb{Z}
\qquad\text{and}\qquad
b\neq 0.
\]
Because $b\neq 0$ and $e\neq 0$, the integer product $be$ is nonzero:
\[
be\neq 0.
\]
Therefore
\[
(ae,be)
\]
is also an admissible fraction-pair.

\textbf{Step 2.} Reduce equality of equivalence classes to the representative relation.

To prove
\[
[(a,b)] = [(ae,be)],
\]
it is enough to prove that the representatives are equivalent:
\[
(a,b)\sim(ae,be).
\]

\textbf{Step 3.} Apply the definition of the fraction-pair relation.

By the definition of $\sim$,
\[
(a,b)\sim(ae,be)
\]
holds exactly when
\[
a(be)=b(ae).
\]

\textbf{Step 4.} Verify the cross-product equality.

Using associativity of integer multiplication,
\[
a(be)=(ab)e.
\]
Using commutativity of integer multiplication,
\[
(ab)e=(ba)e.
\]
Using associativity again,
\[
(ba)e=b(ae).
\]
Hence
\[
a(be)=b(ae).
\]
Therefore
\[
(a,b)\sim(ae,be).
\]

\textbf{Step 5.} Conclude equality of classes.

Since the representatives are equivalent,
\[
(a,b)\sim(ae,be),
\]
their equivalence classes are equal:
\[
[(a,b)] = [(ae,be)].
\]
\end{proof}

\begin{remark*}[Proof structure]
The proof is definitional. Scaling the numerator and denominator by the
same nonzero integer produces another admissible representative. Equality
of equivalence classes is then reduced to the fraction-pair relation, and
the required cross-product equality follows from associativity and
commutativity of integer multiplication.
\end{remark*}

\begin{dependencies}
\begin{itemize}
  \item \hyperref[def:fraction-pair-equality]{Definition~\ref*{def:fraction-pair-equality}}
  \item Commutativity and associativity of integer multiplication
  \item Nonzero product property for integers
\end{itemize}
\end{dependencies}



