% =========================================================
% Rational Numbers
% =========================================================
\chapter{Rational Numbers ($\mathbb{Q}$)}

\begin{tcolorbox}[
  colback=breadcrumb!8,
  colframe=breadcrumb!50,
  arc=2pt,
  left=8pt, right=8pt, top=6pt, bottom=6pt,
  title={\small\textbf{rationals}},
  fonttitle=\small\bfseries
]
\begin{center}
\small
Integers ($\mathbb{Z}$)
$\;\to\;$ \textbf{Rationals ($\mathbb{Q}$)}
$\;\to\;$ Real Numbers ($\mathbb{R}$)
\end{center}
\end{tcolorbox}

\begin{tcolorbox}[
  colback=gray!6,
  colframe=gray!40,
  arc=2pt,
  left=8pt, right=8pt, top=6pt, bottom=6pt,
  title={\small\textbf{Structural Roadmap}},
  fonttitle=\small\bfseries
]
\noindent\textbf{How we got here.}
The integers gave us additive inverses, so subtraction became internal to the number system.
Division still fails: for example, $1 \div 2$ is not an integer.  The rational numbers adjoin
multiplicative inverses for all nonzero integers.

\medskip
\noindent\textbf{What this chapter develops.}
We construct $\mathbb{Q}$ from $\mathbb{Z}$ using equivalence classes of ordered pairs, prove that
$\mathbb{Q}$ is an ordered field, study its density and approximation properties, identify the
problems that $\mathbb{Q}$ cannot resolve, and then see why $\mathbb{Q}$ is still incomplete.

\medskip
\noindent\textbf{Where this leads.}
The rational numbers are already rich enough to support most of arithmetic and much of algebra, but
not rich enough to capture all limits.  That failure motivates the construction of $\mathbb{R}$.
\end{tcolorbox}


\input{volume-ii/rationals/notes/index}

\input{volume-ii/rationals/proofs/index}


\section*{Capstone}
\input{volume-ii/rationals/proofs/exercises/index}
