\newpage
\phantomsection
\label{prf:iterator-relation-consistency}
\LRAProofFor{lem:iterator-relation-consistency}

\begin{remark*}[Return]
\hyperref[lem:iterator-relation-consistency]{Return to Lemma}
\end{remark*}

\begin{remark*}[Proof status]
\ProofVaultURL{https://github.com/wsollers/lra-proof-vault/tree/master/volume-ii/peano-systems/lem-iterator-relation-consistency}
\end{remark*}

\begin{lemma*}[Consistency of the Minimal Iterator Relation]
Let $(P,S,1)$ be a Peano system, let $(W,c,g)$ be iterator data for it, and let
$R^*$ be the minimal iterator relation for $(W,c,g)$. Then $R^*$ is an
iterator relation for $(W,c,g)$.
\end{lemma*}

\begin{proof}[Professional Standard Proof]
\LRAProofBodyStart
Let
\[
\mathcal R=
\{R\subseteq P\times W : R \text{ is an iterator relation for } (W,c,g)\}.
\]
By definition,
\[
R^*=\bigcap \mathcal R.
\]
We verify the three clauses in the definition of an iterator relation.

First, since every member of $\mathcal R$ is a subset of $P\times W$, its
intersection is also a subset of $P\times W$. Hence
\[
R^*\subseteq P\times W.
\]

Second, let $R\in\mathcal R$. Because $R$ is an iterator relation,
$(1,c)\in R$. Since this holds for every $R\in\mathcal R$, the pair $(1,c)$
belongs to the intersection. Thus
\[
(1,c)\in R^*.
\]

Third, suppose $n\in P$, $w\in W$, and $(n,w)\in R^*$. Let $R\in\mathcal R$.
Then $(n,w)\in R$, because $R^*$ is the intersection of all relations in
\(\mathcal R\). Since $R$ is an iterator relation, it is closed under the
iterator step, so
\[
(S(n),g(w))\in R.
\]
This holds for every $R\in\mathcal R$, and therefore
\[
(S(n),g(w))\in R^*.
\]

We have shown that $R^*\subseteq P\times W$, that $(1,c)\in R^*$, and that
$R^*$ is closed under the iterator step. Hence $R^*$ is an iterator relation
for $(W,c,g)$.
\end{proof}

\begin{proof}[Detailed Learning Proof]
\LRAProofBodyStart

\textbf{Step 1. Identify the family being intersected.}
Let
\[
\mathcal R=
\{R\subseteq P\times W : R \text{ is an iterator relation for } (W,c,g)\}.
\]
By definition, the minimal iterator relation is
\[
R^*=\bigcap \mathcal R.
\]
To prove that $R^*$ is itself an iterator relation, each clause in the
definition of iterator relation must be checked.

\textbf{Step 2. Check that $R^*$ is a relation from $P$ to $W$.}
Every member of $\mathcal R$ is a subset of $P\times W$. An intersection of
subsets of $P\times W$ is again a subset of $P\times W$. Therefore
\[
R^*\subseteq P\times W.
\]

\textbf{Step 3. Check the base pair.}
Let $R\in\mathcal R$. Since $R$ is an iterator relation for $(W,c,g)$, it
contains the base pair $(1,c)$. This is true for every $R\in\mathcal R$.
Hence $(1,c)$ belongs to the intersection of all such relations, so
\[
(1,c)\in R^*.
\]

\textbf{Step 4. Check closure under the iterator step.}
Suppose $n\in P$, $w\in W$, and $(n,w)\in R^*$. Let $R\in\mathcal R$ be
arbitrary. Since $R^*$ is the intersection of all relations in $\mathcal R$,
the membership $(n,w)\in R^*$ implies $(n,w)\in R$.

Because $R$ is an iterator relation, it is closed under the iterator step.
Therefore
\[
(S(n),g(w))\in R.
\]
This holds for every $R\in\mathcal R$. Hence $(S(n),g(w))$ belongs to the
intersection of all relations in $\mathcal R$, and so
\[
(S(n),g(w))\in R^*.
\]

\textbf{Step 5. Conclude.}
The relation $R^*$ is a subset of $P\times W$, contains the base pair
$(1,c)$, and is closed under the iterator step. These are exactly the clauses
in the definition of an iterator relation. Hence $R^*$ is an iterator relation
for $(W,c,g)$.
\end{proof}

\begin{remark*}[Proof structure]
The proof uses the closure properties that define the class of iterator
relations and verifies that those properties are preserved by intersection.
The base-pair and step-closure clauses hold in every iterator relation in the
family, so they hold in the intersection that defines $R^*$. The argument is
therefore a direct verification from the definition of the minimal iterator
relation.
\end{remark*}

\begin{dependencies}
\begin{itemize}
  \item \hyperref[def:peano-system]{Peano System}
  \item \hyperref[def:iterator-data-on-peano-system]{Iterator Data on a Peano System}
  \item \hyperref[def:iterator-relation]{Iterator Relation}
  \item \hyperref[def:minimal-iterator-relation]{Minimal Iterator Relation}
\end{itemize}
\end{dependencies}

\clearpage
