% =========================================================
% Farey Sequences, Stern--Brocot, and Limiting Processes
% =========================================================

\subsection{Farey Sequences, Stern--Brocot, and Limiting Processes}
\label{sec:sb-limiting-processes}

The mediant operation, Farey sequences, and the Stern--Brocot tree are
often introduced as combinatorial or number-theoretic constructions on
the rational numbers.  However, they also encode genuine limiting
processes.  In this sense they form a bridge between number theory and
analysis.

The key idea is that repeated rational refinement produces sequences of
rationals converging to real numbers.  Thus these constructions do not
merely organize $\mathbb{Q}$; they also point beyond $\mathbb{Q}$ toward
$\mathbb{R}$.

\begin{remark}
There are at least three major ways in which Farey and Stern--Brocot
structures enter limiting processes:

\begin{enumerate}
\item they generate convergent sequences of rational approximations;
\item they control approximation errors through exact gap formulas;
\item they encode irrational numbers as limits of infinite refinement
paths.
\end{enumerate}

\end{remark}

% ---------------------------------------------------------
% Stern--Brocot paths as limiting processes
% ---------------------------------------------------------

\begin{theorem}[Stern--Brocot interval refinement]
\label{thm:stern-brocot-refinement}

Let $x>0$ be a real number.  Suppose one begins with two rational
numbers

\[
\frac{a}{b} < x < \frac{c}{d},
\qquad b,d>0,
\]

and repeatedly inserts the mediant

\[
\frac{a+c}{b+d}.
\]

At each step, replace one endpoint by the mediant so that the new
interval still contains $x$.  Then this process generates a nested
sequence of rational intervals containing $x$.
\hyperref[prf:stern-brocot-refinement]{Go to proof.}
\end{theorem}

\begin{remark}
This is a rational form of interval bisection.  Instead of dividing an
interval at its midpoint, one divides it at its mediant.
\end{remark}

\begin{remark}
If the process is continued indefinitely, one obtains a sequence of
nested intervals

\[
\left[\frac{a_1}{b_1},\frac{c_1}{d_1}\right]
\supseteq
\left[\frac{a_2}{b_2},\frac{c_2}{d_2}\right]
\supseteq
\left[\frac{a_3}{b_3},\frac{c_3}{d_3}\right]
\supseteq \cdots
\]

whose endpoints are rational numbers and whose lengths shrink.
This is the first point at which the Stern--Brocot picture begins to
look like an analytic limiting process rather than a purely arithmetic
construction.
\end{remark}

% ---------------------------------------------------------
% Mediant-generated convergent sequences
% ---------------------------------------------------------

\begin{remark}[Mediants generate convergent sequences]
Starting with two rational bounds and repeatedly inserting mediants
often produces monotone sequences of rational numbers converging to a
real number.

For example, to approximate $\sqrt{2}$ one may begin with

\[
1 < \sqrt{2} < 2.
\]

Repeated refinement yields rational approximations such as

\[
\frac{3}{2},\quad
\frac{7}{5},\quad
\frac{17}{12},\quad
\frac{41}{29},\dots
\]

which converge to $\sqrt{2}$.

Thus the mediant process is not only a way to find intermediate
rationals; it is also a mechanism for constructing convergent rational
sequences.
\end{remark}

% ---------------------------------------------------------
% Farey gap and approximation
% ---------------------------------------------------------

\begin{remark}[Farey neighbors control approximation error]
If

\[
\frac{a}{b} < \frac{c}{d}
\]

are Farey neighbors, then

\[
bc-ad=1
\]

and hence

\[
\frac{c}{d}-\frac{a}{b}=\frac{1}{bd}.
\]

Therefore, if a real number $x$ lies between these two rationals, then
its distance from either endpoint is controlled by a quantity of order

\[
\frac{1}{bd}.
\]

As the denominators grow, these gaps shrink.  Thus the Farey gap formula
gives a precise arithmetic explanation of why rational approximations
can become arbitrarily accurate.
\end{remark}

% ---------------------------------------------------------
% Dirichlet link
% ---------------------------------------------------------

\begin{remark}[Connection with Dirichlet approximation]
Dirichlet's approximation theorem states that for every real number $x$
there exist infinitely many rational numbers $\frac{p}{q}$ such that

\[
\left|x-\frac{p}{q}\right| < \frac{1}{q^2}.
\]

This bound is much stronger than mere density.  Its geometric content is
that one can find rational approximations inside intervals whose lengths
shrink roughly like reciprocal quadratic functions of the denominator.

Farey gaps and Stern--Brocot refinement provide a concrete geometric
picture of how such small intervals arise.
\end{remark}

% ---------------------------------------------------------
% Continued fraction bridge
% ---------------------------------------------------------

\begin{remark}[Continued fractions and Stern--Brocot]
The Stern--Brocot tree and continued fractions encode the same basic
approximation process.

Every irrational number has an infinite continued fraction expansion

\[
x
=
a_0 +
\cfrac{1}{a_1 +
\cfrac{1}{a_2 +
\cfrac{1}{a_3 + \cdots}}}.
\]

Its convergents

\[
\frac{p_0}{q_0},\;
\frac{p_1}{q_1},\;
\frac{p_2}{q_2},\dots
\]

satisfy

\[
\lim_{n\to\infty}\frac{p_n}{q_n}=x.
\]

These same approximants can be interpreted as arising from a path in
the Stern--Brocot tree.  Thus the Stern--Brocot tree gives a
combinatorial model of continued-fraction convergence.
\end{remark}

% ---------------------------------------------------------
% The central limit theorem for Stern--Brocot
% ---------------------------------------------------------

\begin{theorem}[Stern--Brocot Limit Theorem]
\label{thm:stern-brocot-limit}

Every irrational number $x>0$ determines a unique infinite path in the
Stern--Brocot tree.  The sequence of rational numbers encountered along
this path converges to $x$.

Conversely, every infinite path in the Stern--Brocot tree determines a
unique positive irrational number as its limit.
\hyperref[prf:stern-brocot-limit]{Go to proof.}
\end{theorem}

\begin{remark}[Interpretation]
This theorem says that the Stern--Brocot tree does not merely enumerate
the rational numbers.  Its infinite paths encode all positive irrational
numbers as limits of rational approximations.

Thus:

\begin{itemize}
\item finite Stern--Brocot paths correspond to rational numbers;
\item infinite Stern--Brocot paths correspond to irrational numbers.
\end{itemize}

In this way the Stern--Brocot tree reaches beyond $\mathbb{Q}$ and
touches the full continuum of positive real numbers.
\end{remark}

\begin{remark}[Why this matters in analysis]
This is one of the first places in elementary analysis where every
irrational number appears through a single systematic construction based
entirely on rational numbers.

The slogan is:

\[
\text{infinite rational refinement} \longrightarrow \text{irrational limit}.
\]

That makes the theorem especially important conceptually.  It shows that
irrational numbers are not mysterious isolated objects added from
outside; they arise naturally as limits of coherent rational processes.
\end{remark}

% ---------------------------------------------------------
% Nested intervals viewpoint
% ---------------------------------------------------------

\begin{remark}[Nested interval viewpoint]
The Stern--Brocot limit process can also be understood through nested
intervals.

Each step produces a smaller rational interval containing the target
number.  Thus one obtains

\[
I_1 \supseteq I_2 \supseteq I_3 \supseteq \cdots
\]

with

\[
x \in I_n
\quad\text{for all }n,
\]

and with lengths tending to $0$.

Hence the limit point is singled out by an infinite nested system of
rational intervals.  This is closely related in spirit to the nested
interval principle and to Dedekind's idea of locating a real number by
order cuts.
\end{remark}

% ---------------------------------------------------------
% Conceptual synthesis
% ---------------------------------------------------------

\subsubsection{Consequences}

\begin{remark}[Why the chapter forces the issue]
The theorem flow of this chapter does not merely show that rational
numbers are useful. It shows that rational methods inevitably generate
processes whose natural endpoints are not always rational.

That pressure builds in a very specific order:
\begin{enumerate}
\item mediant refinement and Stern--Brocot paths generate nested
      rational intervals;
\item Farey gaps explain why those intervals become arbitrarily small;
\item rational approximation theorems show that the target can be
      approached as closely as one likes;
\item Cauchy-style limiting behavior appears naturally inside
      $\mathbb{Q}$;
\item yet some of the targets produced by these processes are missing
      from $\mathbb{Q}$ itself.
\end{enumerate}
\end{remark}

\begin{remark}[The first visible obstruction: the gap at $\sqrt{2}$]
The most concrete example is
\hyperref[thm:sqrt2-irrational]{Theorem~\ref*{thm:sqrt2-irrational}}:
there is no rational number whose square is $2$.

This is not an isolated curiosity. By
\hyperref[lem:rational-square-bracketing]{Lemma~\ref*{lem:rational-square-bracketing}}
and
\hyperref[thm:rational-square-approximation-from-above]{Theorem~\ref*{thm:rational-square-approximation-from-above}},
positive rational numbers can be trapped between rational squares and
approached by rational squares with arbitrarily small error. So the
arithmetic of $\mathbb{Q}$ naturally produces sequences that home in on
locations that $\mathbb{Q}$ itself does not occupy.
\end{remark}

\begin{remark}[Refinement without completion]
The mediant and Stern--Brocot constructions sharpen this picture.
By
\hyperref[thm:stern-brocot-refinement]{Theorem~\ref*{thm:stern-brocot-refinement}},
one obtains nested rational intervals containing a target real number.
By
\hyperref[thm:rational-chain-with-small-steps]{Theorem~\ref*{thm:rational-chain-with-small-steps}},
rational intervals can also be traversed by arbitrarily small rational
increments. These are exactly the kinds of structures that suggest
limit formation.

But the chapter's point is that repeated rational refinement does not
guarantee a rational endpoint. The process is internally coherent while
its limit may lie outside $\mathbb{Q}$.
\end{remark}

\begin{remark}[Cauchy pressure]
This is the same tension that appears in the study of rational Cauchy
sequences. A sequence can become arbitrarily tightly packed in the
internal metric sense and still fail to converge to a rational number.
The problem with $\mathbb{Q}$ is therefore not lack of density or lack
of approximation power. The problem is failure of completion.
\end{remark}

\begin{remark}[The inescapable conclusion]
Taken together, the theorem flow of this chapter leads to a single
conclusion:
\[
\text{rational refinement}
\;\longrightarrow\;
\text{rational approximation}
\;\longrightarrow\;
\text{Cauchy behavior}
\;\not\Longrightarrow\;
\text{rational limit}.
\]

That is the precise sense in which $\mathbb{Q}$ is incomplete.
The real numbers are introduced not because the rationals are too
sparse, but because $\mathbb{Q}$ does not contain all the limits
naturally produced by its own approximation processes.
\end{remark}

\begin{remark}[Conceptual synthesis]
Farey sequences, mediants, Stern--Brocot refinement, continued
fractions, and Dirichlet approximation are not isolated topics.  They
form a single chain of ideas:

\[
\text{Farey neighbors}
\;\longrightarrow\;
\text{mediant refinement}
\;\longrightarrow\;
\text{Stern--Brocot tree}
\;\longrightarrow\;
\text{continued fractions}
\;\longrightarrow\;
\text{best rational approximations}
\;\longrightarrow\;
\text{limits in }\mathbb{R}.
\]

This chain shows that the rational numbers already contain, in
compressed form, the structure needed to approximate every real number.
\end{remark}

\begin{remark}[Three viewpoints on completing $\mathbb{Q}$]
The passage from $\mathbb{Q}$ to $\mathbb{R}$ can be understood from
three complementary perspectives:

\begin{enumerate}
\item \textbf{Dedekind cuts:} real numbers are determined by order
structure;
\item \textbf{Cauchy sequences:} real numbers are determined by metric
completion;
\item \textbf{Stern--Brocot paths:} real numbers are determined by
infinite rational approximation processes.
\end{enumerate}

These are not three different continua.  They are three different ways
of describing the same completed number system.
\end{remark}

\begin{remark}[A foundational perspective]
The rational numbers $\mathbb{Q}$ are dense, ordered, Archimedean, and
arithmetically rich.  Yet they are incomplete.

The Stern--Brocot Limit Theorem makes this incompleteness vivid in a
particularly concrete way.  It shows that every irrational number can be
approached by an infinite, structured, purely rational process.

Thus one may think of the real numbers as arising when one allows not
only rational numbers themselves, but also the limits of all coherent
rational refinement procedures.

Symbolically:

\[
\mathbb{Q}
\quad+\quad
\text{all limiting rational processes}
\quad=\quad
\mathbb{R}.
\]

\end{remark}
