\newpage
\phantomsection
\label{prf:one-is-unique-non-successor}

\begin{remark*}[Return]
\hyperref[cor:one-is-unique-non-successor]{Return to Corollary}
\end{remark*}

\begin{corollary*}[The Distinguished Element Is the Unique Non-Successor]
Let $(P,S,1)$ be a Peano system. An element $x \in P$ is not a successor of
any element of $P$ if and only if $x=1$.
\end{corollary*}

\begin{proof}
\textbf{Professional Standard Proof.}~\\
Assume first that $x\in P$ is not a successor of any element of $P$. By
\hyperref[thm:every-element-is-one-or-a-successor]{Every Element Is Either
$1$ or a Successor}, either $x=1$ or there exists $u\in P$ such that
$S(u)=x$. The second alternative contradicts the assumption that $x$ is not a
successor. Hence $x=1$.

Conversely, suppose $x=1$. By the Peano axiom that $1$ is not a successor,
there is no $u\in P$ such that $S(u)=1$. Since $x=1$, there is no $u\in P$
such that $S(u)=x$. Therefore $x$ is not a successor of any element of $P$.
\end{proof}

\begin{proof}
\textbf{Detailed Learning Proof.}~\\
We prove both implications.

First suppose $x$ is not a successor of any element of $P$. Since $x\in P$,
the dichotomy theorem says that $x$ has one of two forms:
\[
x=1
\]
or
\[
\exists u\in P\;(S(u)=x).
\]
The second possibility says exactly that $x$ is a successor of an element of
$P$, which is excluded by hypothesis. Therefore the only remaining
possibility is $x=1$.

Conversely suppose $x=1$. The Peano axiom
\hyperref[ax:peano-one-not-successor]{Distinguished Element Is Not a
Successor} says that no element of $P$ has successor equal to $1$. Hence no
element of $P$ has successor equal to $x$. Therefore $x$ is not a successor
of any element of $P$.
\end{proof}

\begin{remark*}[Proof structure]
The forward implication uses the dichotomy theorem and eliminates the
successor case. The reverse implication is exactly the Peano axiom that the
distinguished element is not a successor.
\end{remark*}

\begin{remark*}[Dependencies]~\\
\begin{itemize}
  \item \hyperref[ax:peano-one-not-successor]{Distinguished Element Is Not a Successor}
  \item \hyperref[thm:every-element-is-one-or-a-successor]{Every Element Is Either $1$ or a Successor}
\end{itemize}
\end{remark*}

