\section{Algebraic Structure of \texorpdfstring{$\mathbb{N}$}{N}}
\label{sec:algebraic-structure-of-n}

\begin{topicbox}{$\mathbb{N}$ as a Commutative Monoid Under Addition}
\begin{exposition}
The arithmetic established in the preceding sections can now be packaged into
a named algebraic structure. Under addition, $\mathbb{N}$ is a commutative
monoid with identity $0$ --- or, in the Peano-system setting that uses $1$
as the distinguished element with no additive identity available internally,
the correct statement is that $(\mathbb{N}, +)$ satisfies associativity and
commutativity. This topic states the structural conclusion that follows from
the individual arithmetic theorems proved above.
\end{exposition}

\begin{tcolorbox}[colback=thmbox, colframe=thmborder, arc=2pt,
  left=6pt, right=6pt, top=4pt, bottom=4pt,
  title={\small\textbf{Theorem ($\mathbb{N}$ Is Closed, Associative, and Commutative Under Addition)}},
  fonttitle=\small\bfseries]
\begin{theorem}[$\mathbb{N}$ Is Closed, Associative, and Commutative Under Addition]
\label{thm:n-additive-structure}
Let $(P,S,1)$ be a Peano system. The binary operation $+$ on $P$ satisfies:
\begin{enumerate}
  \item \textbf{Closure.} For all $a,b \in P$, $a+b \in P$.
  \item \textbf{Associativity.} For all $a,b,c \in P$,
  $(a+b)+c = a+(b+c)$.
  \item \textbf{Commutativity.} For all $a,b \in P$, $a+b = b+a$.
  \item \textbf{Left cancellation.} For all $a,b,c \in P$, if $a+b = a+c$ then $b=c$.
  \item \textbf{Right cancellation.} For all $a,b,c \in P$, if $b+a = c+a$ then $b=c$.
\end{enumerate}
\hyperref[prf:n-additive-structure]{Go to proof.}
\end{theorem}
\end{tcolorbox}

\begin{remark*}[Standard quantified statement]
\[
\begin{aligned}
&\text{Let } (P,S,1) \text{ be a Peano system.}\\
&\forall a,b,c \in P:\;
(a+b)+c = a+(b+c)
\;\land\;
a+b = b+a
\;\land\;
(a+b=a+c \Longrightarrow b=c).
\end{aligned}
\]
\end{remark*}

\begin{remark*}[Interpretation]
This theorem is a packaging result. It names the algebraic structure that
has been proved piece by piece: the three individual theorems on associativity,
commutativity, and cancellation together certify that $(P,+)$ is a cancellative
commutative semigroup. There is no additive identity in $P$ because $P$ has
no element $0$ with $0+n=n$ for all $n$; the smallest element under addition
is $1$, not an identity.
\end{remark*}

\begin{remark*}[Dependencies]
\begin{itemize}
  \item \hyperref[thm:addition-associative]{Addition Is Associative}
  \item \hyperref[thm:addition-commutative]{Addition Is Commutative}
  \item \hyperref[thm:addition-left-cancellation]{Left Cancellation for Addition}
  \item \hyperref[thm:addition-right-cancellation]{Right Cancellation for Addition}
\end{itemize}
\end{remark*}
\end{topicbox}

\begin{topicbox}{$\mathbb{N}$ as a Commutative Monoid Under Multiplication}
\begin{exposition}
Multiplication on $P$ has a two-sided multiplicative identity, namely $1$,
the distinguished first element. Under multiplication, $P$ is therefore a
commutative monoid. This topic states the structural packaging for multiplication
that parallels the additive packaging above.
\end{exposition}

\begin{tcolorbox}[colback=thmbox, colframe=thmborder, arc=2pt,
  left=6pt, right=6pt, top=4pt, bottom=4pt,
  title={\small\textbf{Theorem ($\mathbb{N}$ Is a Commutative Monoid Under Multiplication)}},
  fonttitle=\small\bfseries]
\begin{theorem}[$\mathbb{N}$ Is a Commutative Monoid Under Multiplication]
\label{thm:n-multiplicative-monoid}
Let $(P,S,1)$ be a Peano system. The binary operation $\cdot$ on $P$ satisfies:
\begin{enumerate}
  \item \textbf{Closure.} For all $a,b \in P$, $a \cdot b \in P$.
  \item \textbf{Associativity.} For all $a,b,c \in P$,
  $(a \cdot b) \cdot c = a \cdot (b \cdot c)$.
  \item \textbf{Commutativity.} For all $a,b \in P$,
  $a \cdot b = b \cdot a$.
  \item \textbf{Identity.} For all $a \in P$,
  $1 \cdot a = a = a \cdot 1$.
\end{enumerate}
Consequently $(P, \cdot, 1)$ is a commutative monoid.
\hyperref[prf:n-multiplicative-monoid]{Go to proof.}
\end{theorem}
\end{tcolorbox}

\begin{remark*}[Standard quantified statement]
\[
\begin{aligned}
&\text{Let } (P,S,1) \text{ be a Peano system.}\\
&\forall a,b,c \in P:\;
(a \cdot b) \cdot c = a \cdot (b \cdot c)
\;\land\;
a \cdot b = b \cdot a
\;\land\;
1 \cdot a = a.
\end{aligned}
\]
\end{remark*}

\begin{remark*}[Interpretation]
The distinguished first element $1$ is a multiplicative identity on both
sides, so multiplication promotes $P$ from a semigroup to a monoid. The
absence of multiplicative inverses (other than $1^{-1} = 1$, which does not
lie in $P$ in the usual sense) is what prevents $(P,\cdot)$ from being a
group. The structural limitation is intrinsic: no element of $P$ other than
$1$ has a reciprocal in $P$.
\end{remark*}

\begin{remark*}[Dependencies]
\begin{itemize}
  \item \hyperref[thm:multiplication-associative]{Multiplication Is Associative}
  \item \hyperref[thm:multiplication-commutative]{Multiplication Is Commutative}
  \item \hyperref[thm:multiplication-with-one]{Multiplication With $1$}
  \item \hyperref[thm:one-times-n]{One Times $n$}
\end{itemize}
\end{remark*}
\end{topicbox}

\begin{topicbox}{$\mathbb{N}$ as a Semiring and the Absence of Subtraction}
\begin{exposition}
Together, addition and multiplication on $P$ form a semiring: addition is a
commutative semigroup, multiplication is a commutative monoid, and
multiplication distributes over addition on both sides. The critical structural
limitation is that subtraction is not a total operation on $P$. Since $P$
has no additive identity and no additive inverses, the equation $a + x = b$
has a solution in $P$ only when $b > a$; it has no solution when $b \le a$.
This failure of internal subtraction is the reason $\mathbb{Z}$ is introduced
immediately after $\mathbb{N}$.
\end{exposition}

\begin{tcolorbox}[colback=thmbox, colframe=thmborder, arc=2pt,
  left=6pt, right=6pt, top=4pt, bottom=4pt,
  title={\small\textbf{Theorem ($\mathbb{N}$ Is a Semiring)}},
  fonttitle=\small\bfseries]
\begin{theorem}[$\mathbb{N}$ Is a Semiring]
\label{thm:n-is-semiring}
Let $(P,S,1)$ be a Peano system. Then $(P,+,\cdot)$ satisfies:
\begin{enumerate}
  \item $(P,+)$ is a commutative semigroup.
  \item $(P,\cdot,1)$ is a commutative monoid.
  \item Multiplication distributes over addition: for all $a,b,c \in P$,
  \[
  a \cdot (b+c) = (a \cdot b) + (a \cdot c).
  \]
\end{enumerate}
\hyperref[prf:n-is-semiring]{Go to proof.}
\end{theorem}
\end{tcolorbox}

\begin{remark*}[Standard quantified statement]
\[
\begin{aligned}
&\text{Let } (P,S,1) \text{ be a Peano system.}\\
&\forall a,b,c \in P:\;
a \cdot (b+c) = (a \cdot b) + (a \cdot c)
\;\land\;
(a+b)+c = a+(b+c)
\;\land\;
a+b = b+a
\;\land\;
1 \cdot a = a.
\end{aligned}
\]
\end{remark*}

\begin{remark*}[Interpretation]
A semiring packages the interaction of two operations without requiring
additive inverses. The natural numbers are the canonical example of a semiring
that is not a ring: they satisfy every ring axiom except the existence of
additive inverses, which fails because there is no element $-1$ in $P$.
The integers are the smallest ring extending $P$ that repairs this failure.
\end{remark*}

\begin{remark*}[Dependencies]
\begin{itemize}
  \item \hyperref[thm:n-additive-structure]{$\mathbb{N}$ Is Closed, Associative, and Commutative Under Addition}
  \item \hyperref[thm:n-multiplicative-monoid]{$\mathbb{N}$ Is a Commutative Monoid Under Multiplication}
  \item \hyperref[thm:multiplication-distributes-over-addition]{Multiplication Distributes Over Addition}
\end{itemize}
\end{remark*}
\end{topicbox}
